---
title: "Class 19 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook: 
    df_print: paged
    highlight: haddock
    theme: readable
---

## Learning Objective
Learn to map functions to elements in a vector

## Topics
1. Learn the basic patterns of mapping
2. Use it to clean the UN migration data from here: https://www.un.org/en/development/desa/population/migration/data/estimates2/estimates17.asp


```{r}
plot_hist <- function(num_vec) {
    if (is.numeric(num_vec)) hist(num_vec) else print("I am a character")
}
```

```{r}
library(gapminder)
gapminder
```

```{r}
map(gapminder, plot_hist)
```


```{r}
glimpse(gapminder)
```

NON RECOMMENDED WAY TO DO THIS
```{r}
for (i in gapminder) {
    # print(i)
    plot_hist(i)
}
```

```{r}
seq_along(gapminder)


```

RECOMMENDED WAY
```{r}
for (i in seq_along(gapminder)) {
    # print(i)
    plot_hist(gapminder[[i]])
}
```


## Task

1. I want to iterate through each sheet and run a function that automatically cleans it up


First step is to identify all the sheets in the excel file

```{r}
library(readxl); library(tidyverse); 

list.files("data")
```


```{r}
migrant_sheets <- excel_sheets("data/UN_MigrantStockTotal_2017.xlsx")

migrant_sheets
```


Lets remove all the unwanted sheets.

```{r}
migrant_sheets <- migrant_sheets[str_detect(migrant_sheets, pattern = "Table")]
migrant_sheets <- migrant_sheets[migrant_sheets != "Table 4"]

migrant_sheets
```


Map through each of these sheets and load the data.
```{r}
all_data <- map(migrant_sheets, read_excel, path = "data/UN_MigrantStockTotal_2017.xlsx", .name_repair = janitor::make_clean_names, skip = 15)
```


```{r}
all_data
```


```{r}
clean_data <- function(df) {
  df %>% 
    select(-c(1, 3, 5:12)) %>% 
    filter(x_4 < 900) %>% 
    gather(key = "year", value = "migrant_stock", -(1:2)) %>% 
    separate(col = "year", into = c("year", "gender"), sep = "_") %>% 
    mutate(year = str_replace(string = year, pattern = "x", replacement = ""),
           gender = if_else(gender == "2", "male", "female")) %>% 
    select(-x_4)
}
```


```{r}
map(all_data, clean_data)
```

```{r}
map(migrant_sheets, read_excel, path = "data/UN_MigrantStockTotal_2017.xlsx", .name_repair = janitor::make_clean_names, skip = 15) %>% 
  map(., clean_data)
```


