---
title: "Class 13 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---

**LEARNING OBJECTIVE**: Understand the principles of tidy data

**TOPICS**

1. What is tidy data?
2. Examples of messy data
3. Spread columns to convert a long table to wide
4. Gather columns to convert a wide table to long
5. Separate a column
6. Clean the who data (https://www.who.int/tb/country/data/download/en/)


##What is tidy data?

```{r}
library(tidyverse)

table1
```

```{r}
table2
```

```{r}
table3
```


```{r}
table5
```


##Spread from long to wide

```{r}
table2
```

```{r}
table2 %>% 
    spread(key = type, value = count)
```

```{r}
table1
```


##lets separate a column into two
```{r}
table3 %>% 
    separate(col = rate, into = c("cases", "population"), sep = "/", convert = T)
```

##Lets unite a columns and separate another

```{r}
table5
```

```{r}
table5 %>% 
    unite(col = "year", century, year, sep = "")
```

I can also use a mutate to do this

```{r}
table5 %>% 
    mutate(newYear = paste(century, year, sep = ""))
```

```{r}
table5 %>% 
    unite(col = "year", century, year, sep = "")  %>% 
    separate(col = rate, into = c("cases", "population"))
```



##Lets try again

Lets try to gather multiple columns into a single one.

```{r}
table1
```

I want to gather cases and population into a single column and create an untidy tibble.
```{r}
table1 %>% 
    gather(key = "emmasKey", value = "newValue", cases, population)
```

Another way to specify these columns
```{r}
table1 %>% 
    gather(key = "casePop", value = "value", -country, -year)
```

```{r}
table1
```


```{r}
table1 %>% 
    gather(key = "casePop", value = "value")
```


```{r}
table1 %>% 
    gather(key = "casePop", value = "value", -country)
```


```{r}
weather
```


```{r}
weather %>% 
    select(-origin, -year)
```

##A spread redux

Spread requires me to specifiy the key column that needs to be spread into a wide format along with the name of the column that contains the corresp values for those keys.
```{r}
table2 %>% 
    spread(key = type, value = count)
```

```{r}
table2
```


```{r}
table1 %>% 
    spread()
```

##Lets separate again
```{r}
table3
```
```{r}
table3 %>% 
    bind_rows(., tibble(country = "My Country", year = 2000, rate = "3434/454/23423/32432")) %>% 
    separate(col = rate, into = c("cases", "population", "newCol"), sep = "/")
```


```{r}
table3 %>% 
    separate(col = rate, into = c("cases", "population", "newCol"), sep = "/")
```

Gather and Spread mirror each other. Gather gathers columns into key value pairs. Spread spreads a key value pairs into new columns.

Separate separates a combined columns into multiple. Unite combines multiple columns into a single column.







