---
title: "Assignment 2"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
**Due Date**: Sunday, April 28th

**Keyboard Shortcuts**:

1. Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2. Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3. Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4. Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5. Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

**General Instructions**

1. Create a new notebook (copy parts of the YAML header you are interested in keeping) to work on your assignment.
2. The assignment should be submitted via a shareable github link that points to your notebook (.nb.html) file to me on Slack.
3. Make sure the notebook is neatly formatted, and that all the charts and tables and look good when you preview it.

**Task**: Explore at least 5 different causes for departure and arrival delays (combined - not 5 each). Its not necessary that you find a definitive answer for what causes a delay, instead I want you to showcase your ability to explore the data with a combination of analytical text, tables and charts. The final output should be something you can show a prospective employer or client as proof of you ability to work with data in R.


```{r}
library(tidyverse); library(nycflights13); library(skimr)

planes
```

Hypothesis: Larger planes (as measured by seating) are more frequently delayed

Lets take a look at the planes dataset.
```{r}
planes
```

```{r}
skimr::skim(planes)
```


Lets take a closer look at our explanatory variable.
```{r}
ggplot(data = planes, aes(x = seats)) +
    geom_histogram()
```

```{r}
flights
```

Now lets also look at the dep_delay
```{r}
ggplot(data = flights, aes(x = dep_delay)) +
    geom_histogram()
```
This is the dep delay threshold I am interested in values above this. And some reasons why....
```{r}
depThreshold <- 20
```


Now lets join the planes data with the flights. I am only interested in rows with both dep delay and seating info. I am also only interested in positive delays.
```{r}
seatingData <- inner_join(flights, 
           planes %>% 
               rename(yearMfg = year), 
           by = "tailnum") %>% 
    filter(dep_delay > depThreshold)
    
```

Lets look at a table with some summary stats grouped by seating.
```{r}
seatingData %>% 
    group_by(seats) %>% 
    ##lets remove groups which have less than 20 obs.
    filter(n() > 20) %>% 
    summarise(nObs = n(), 
              meanDelay = mean(dep_delay, na.rm = T)) %>% 
    arrange(desc(meanDelay))
```

This chart is not telling me a compelling story yet.
```{r}
seatingData %>% 
    group_by(seats) %>% 
    ##lets remove groups which have less than 20 obs.
    filter(n() > 20) %>% 
    summarise(nObs = n(), 
              meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = seats, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm")
```


I would like to maybe work with the raw data instead of summarised.

```{r}
seatingData %>% 
    mutate(seatCat = cut_width(seats, width = 100)) %>% 
    ggplot(aes(x = seatCat, y = dep_delay)) +
    geom_jitter(aes(colour = origin), alpha = 0.1) +
    geom_boxplot(alpha = 0.4) +
    coord_flip() +
    scale_y_continuous(limits = c(depThreshold, 250))
```

