---
title: "The Wilderness Society - Email Marketing Data Analysis"
output: html_notebook
---

```{r}
# load libraries
library(tidyverse)
```

```{r}
# load downloaded data
email_data <- read_csv('data/emails_data.csv') %>% 
    rename_all(funs(make.names(.)))

# clean data for charting
email_data_clean <- email_data %>%
  select(-grep('Click|Actions|Notes|Complaints|X34', current_vars())) %>% 
  separate(Date, into = c('Date', 'Time'), sep = ' ') %>% 
  separate(Date, into = c('Year', 'Month', 'Date')) %>% 
  mutate_at(c('Year', 'Month', 'Date', 'Sent', 'Opened', 'Total.Gifts', 'Total.Raised', 'Bounces', 'Unsubscribed'), as.numeric) %>% 
  filter(Sent > 10000 & Total.Gifts > 0)

email_data_clean$Month <- month.abb[email_data_clean$Month]
```


##How are open rates (opens / sends) trending over time?
```{r}
#min and org data for performance
email_data_short <- email_data_clean %>% 
  mutate(openRate = Opened / Sent) %>% 
  group_by(Year, Month) %>% 
  summarize(meanOpenRate = mean(openRate), totalSent = sum(Sent), na.rm = T) %>% 
  arrange(Year, Month)
```


##How does this correspond to overall list size (sends)?

###Open Rate over time
```{r}
#They specifically wanted these titles - I know they are bad
email_data_short %>% 
  ggplot(aes(Month, meanOpenRate, group=factor(Year), color=factor(Year))) +
  geom_line(na.rm = T) +
  geom_point(na.rm = T) +
  labs(title = "Mean Open Rate / Month", x="Month", color="Year") +
  scale_x_discrete(limits = month.abb)
```


###Total Sent over time
```{r}
email_data_short %>% 
  ggplot(aes(Month, totalSent, group=factor(Year), color=factor(Year))) +
  geom_line(na.rm = T) +
  geom_point(na.rm = T) +
  labs(title = "Total Sent / Month", x="Month", y="Total Sent", color="Year") +
  scale_x_discrete(limits = month.abb) +
  ylim(0, 15000000) +
  facet_grid(Year ~ .)
```


##What were the top 3 emails in the last year by (for example): open rate, conversion rate (gifts / opens), revenue per 1000 recipients (sends), average gift (revenue / donations). Look at this outside of EOY (i.e. excluding mid Nov-December).
```{r}
#reorg data for performance in new subtask
email_data_1 <- email_data_clean %>% 
  filter(Month != 'Dec' & Month != 'Nov') %>% 
  select(1:3, Opened, Sent, Campaign, Total.Gifts, Total.Raised) %>% 
  mutate(openRate = (Opened / Sent), conversionRate = (Total.Gifts / Opened), averageGift = (Total.Raised / Total.Gifts))
```


###Top 3: open rate
```{r}
top_n(email_data_1, 3, openRate)
```


###Top 3: conversion rate (gifts / opens)
```{r}
top_n(email_data_1, 3, conversionRate)
```


###Top 3: average gift (revenue / donations)
```{r}
top_n(email_data_1, 3, averageGift)
```


###Top 3: revenue per 1000 recipients (sends)
```{r}
email_data_1 %>% 
  mutate(Revenue.PerK = Total.Raised / (Sent / 1000)) %>% 
  top_n(., 3, Revenue.PerK)
```


##How do bounce and/or opt out rates (per send) correlate with list size?
```{r}
email_data_2 <- email_data_clean %>% 
  filter(Month != 'Dec' & Month != 'Nov') %>% 
  select(1:3, Opened, Sent, Campaign, Total.Gifts, Total.Raised, Bounces, Unsubscribed) %>% 
  mutate(openRate = (Opened / Sent), conversionRate = (Total.Gifts / Opened), averageGift = (Total.Raised / Total.Gifts), bounceRate = (Bounces / Sent), unsubscribedRate = (Unsubscribed / Opened))

ggplot(email_data_2, aes(Sent, bounceRate)) +
  geom_point(na.rm = T) +
  scale_x_continuous(breaks = pretty(email_data_2$Sent, n = 10)) +
  labs(title = "List Size is a poor predictor of Bounce Rate", x="List Size (Sent)", y="Bounce Rate")
```

```{r}
ggplot(email_data_2, aes(Sent, unsubscribedRate)) +
  geom_point(na.rm = T) +
  scale_x_continuous(breaks = pretty(email_data_2$Sent, n = 15)) +
  labs(title = "", x="List Size (Sent)", y="Opt Out Rate")
```


##Which 3 emails have the highest opt-out rate (Unsubscribe)?
```{r}
email_data_new <- email_data_clean %>%
  filter(Year > 2017) %>% 
  mutate(openRate = (Opened / Sent)) %>% 
  top_n(., 3, Unsubscribed)

print(email_data_new)
```




##IGNORE
```{r eval=FALSE}
## just playing around with http request functionalities
library(httr)
library(jsonlite)

## helper functions

# format httr GET request as text and then parse json to table format
json_parse <- function(req) {
  text <- content(req, 'text', 'UTF-8')
  if (identical(text, '')) warn('No output to parse.')
  fromJSON(text);
}

ships_req <- GET('https://swapi.co/api/starships')
ships_req_parsed <- json_parse(ships_req)
ships <- ships_req_parsed$results
names(ships)
```