---
title: "Class 5 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
**LEARNING OBJECTIVE**: Use *ggplot2* along with its *extensions* to make charts that are * work ready*.

**Topics to cover**

0. Recap + Factor correction
1. ~~Make simple charts~~
    + ~~Scatter plot~~
    + ~~Line charts~~
    + ~~Column chart~~
2. ~~Make facetted plots~~
3. ~~Draw multiple geoms in one plot~~



##Recap
DONT USE THE LEVELS FUNCTION TO SET levels.

##Make simple charts
Lets scatter
```{r, message=FALSE}
library(tidyverse)
library(nycflights13)

hourlyDelay <- flights %>%
  group_by(hour) %>%
  summarise(meanDelay = mean(dep_delay, na.rm = T))

ggplot(data = hourlyDelay, mapping = aes(x = hour, y = meanDelay)) +
  geom_point()
```
Just checking the data to see if the hour variable was calculated correctly in the flights dataset.
```{r}
flights %>%
  mutate(newHour = lubridate::hour(time_hour)) %>%
  filter(newHour == 1)
```

```{r}
ggplot(hourlyDelay, aes(x = hour, y = meanDelay)) +
  geom_line()
```

Lets make a column chart (which in the real world is a bar chart).
```{r}
ggplot(hourlyDelay, aes(x = hour, y = meanDelay)) +
  geom_col()
```

Now lets look at a histogram.

```{r}
ggplot(flights, aes(arr_delay)) +
  geom_histogram(bins = 200)
```

Make a bar chart.
```{r}
ggplot(flights, aes(dest)) +
  geom_bar()
```


##Lets facet

```{r}
hourlyOriginDelay <- flights %>%
  group_by(hour, origin) %>%
  summarise(meanDelay = mean(dep_delay, na.rm = T))
hourlyOriginDelay
```
Lets make a facetted scatter plot
```{r}
ggplot(hourlyOriginDelay, aes(x = hour, y = meanDelay)) +
  geom_point() +
  facet_wrap(~origin)
```

##Multiple geoms in a single plot

```{r}
ggplot(hourlyOriginDelay, aes(x = hour, y = meanDelay)) +
  geom_point() +
  geom_line() +
  facet_wrap(~origin)
```


```{r}
ggplot(hourlyOriginDelay, aes(x = hour, y = meanDelay)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~origin)
```



