---
title: "Class 17 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

## **Learning objectives**
1. Write a for loop
2. Learn when to use a for loop and when to vectorize


## **Topics**
1. For loops
2. First function
3. Map

##Illustration of a slightly more complicated for loop
I would like to take gapminder data set and create a histogram if it is a continous variable and bar chart for  continent and do nothing for country.


```{r}
gapminder
```


Step 1 setup the output vector.
```{r}
out_gap <- vector(mode = "list", length = ncol(gapminder))

```


Step 2 setup iterator

```{r}
seq_along(gapminder)
```

Double brackets first.

```{r}
gapminder[1]
```

```{r}
gapminder[1] %>% str()
```

```{r}
gapminder[[1]] %>% str()
```

```{r}
identical(gapminder$country, gapminder[[1]])
```

```{r}
identical(gapminder$country, gapminder[1])
```

```{r}
for (i in seq_along(gapminder)) {
  print(paste("This is the", i, "iteration", "and the mean is", mean(gapminder[[i]])))
}
```

I need an if else statement to control the flow of code.
```{r}
for (i in seq_along(gapminder)) {
  ## pull out ith col from gapminder
  col_data <- gapminder[[i]]
  
  if (is.numeric(col_data)) { ##check if numeric
    print("numeric")
  } else {
    print("non numeric")
  }
}

```


```{r}
is.numeric(gapminder[1])
```

```{r}
for (i in seq_along(gapminder)) {
  ## pull out ith col from gapminder
  col_data <- gapminder[i]
  col_atomic <- gapminder[[i]]
  col_name <- names(gapminder)[i]
  
  if (is.numeric(col_atomic)) { ##check if numeric
    print(paste("The column is", col_name, "its mean is", mean(col_atomic)))
  } else {
    print("non numeric")
  }
}
```

```{r}
names(gapminder) %>% str()
```


```{r}
for (i in seq_along(gapminder)) {
  ## pull out ith col from gapminder
  col_data <- gapminder[i]
  col_atomic <- gapminder[[i]]
  col_name <- names(gapminder)[[i]]
  
  if (is.numeric(col_atomic)) { ##check if numeric
    ggplot(data = col_data, aes(col_name[i]))
  } else {
    print("non numeric")
  }
}
```

```{r}
colnames(gapminder)
```

```{r}
ggplot(data = gapminder, aes(gdpPercap)) + geom_histogram()
```
















