---
title: "Class 17 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

### Task Set 1 : Put the pieces together
1. Use `vector()` to initiate a numeric vector of length 20
2. Initiate a vector with mode set to list and length that is same as the length of the vector x below.

```{r}
x <- 1:5

```

3. I want a loop that will take different columns (either character or numeric) from a tibble that has 10 columns total and calculate the sum if the column is numeric and create a concatenated character string that combines all the characters if it is a character. What should be the length of the output vector? What should be its mode? Create this output vector.

4. See the gapminder data below.
```{r}
gapminder
```

4.a Use `seq_along()` to identify the indices of all the columns in it. Store it in a variable called `mt_cols`

4.b Compare the length of `mt_cols` with `ncol()`s in gapminder to see that they match.

5. Complete the code below to create a simple for loop

```{r}
for( in c("a", "b", "c", "d")) {
    print(i)
}
```

6. Create a sequence of numbers from 1 to 50 store it in a variable x
```{r}
x <- 1:50
```


7. Complete the code below to write a for loop that returns "even" if a number in the vector you created above is even and "odd" otherwise.
```{r}
odds_evens <- vector("character", length = length(x))
for (i in seq_along(x)) {
    odds_evens[[]] <- if ( %% 2 == 0) { } else { }
}

odds_evens
```

8. Now complete the code below to do the same thing with an `if_else()` statement and store the results in a variable called odds_evens2

```{r}
odds_evens2 <- if_else()

odds_evens2
```

9. Use the `identical()` function to check if the odds_evens is identical to odds_evens2.

### Task Set 2: Make a for loop

1. Take the gapminder dataset. Loop through its columns. If a column is numeric then store the output of the `summary()` command. If it is not numeric, then store NULL. Print the ouput

2. Use `names(gapminder)` to identify the names of columns and set these as the names of the output vector you created.

### Task Set 3: Functions

1. Create a function that takes two numbers (i.e. atomic vectors of length one) and returns the square of whichever number is the greatest. Use `stop("Your error message here")` to throw an error if the input vector is greater than 1 in length. Test your function to make sure it works.

2. Create a function that takes an atomic vector (of any length) as its input. If it numeric then use the `hist()` function from base to plot a histogram else return an error using `stop()`. See the code below for an illustration of the `hist()` command. Test your function to make sure it works.

```{r}
hist(gapminder$lifeExp)
```



