---
title: "Class 17 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

### Task Set 1 : Put the pieces together
1. Use `vector()` to initiate a numeric vector of length 20

```{r}
vector(mode = "numeric", length = 20)
```

2. Initiate a vector with mode set to list and length that is same as the length of the vector x below.

```{r}
x <- 1:5

vector(mode = "list", length = length(x))
```

3. Say, I want to create a loop that will take different columns (either character or numeric) from a tibble that has 10 columns total and calculate the sum if the column is numeric and create a concatenated character string that combines all the characters if it is a character. Don't create this loop but answer the questions below.

What should be the length of the output vector that will store the output from this loop? 

10

What should be its mode? 

list

Initiate this output vector.

```{r}
out_cols <- vector(mode = "list", length = 10)

out_cols
```


4. See the gapminder data below.
```{r}
gapminder
```

```{r}
glimpse(gapminder)
```

4.a Use `seq_along()` to identify the indices of all the columns in it. Store it in a variable called `mt_cols`

Lets try out a few different applications of seq_along()

```{r}
x <- sample(1:100, size = 10)

x
```
If I want the indices of x
```{r}
seq_along(x)
```
```{r}
x[3]
```

```{r}
mt_cols <- seq_along(gapminder)
mt_cols
```

4.b Compare the length of `mt_cols` with `ncol()`s in gapminder to see that they match.
```{r}
ncol(gapminder) == length(mt_cols)
```

5. Complete the code below to create a simple for loop

```{r}
for(i in c("a", "b", "c", "d")) {
    print(i)
}
```

6. Create a sequence of numbers from store it in a variable x

```{r}
x <- sample(1:1000, size = 50)
x
```

```{r}
seq_along(x)
```

```{r}
x[1]
```

7. Complete the code below to write a for loop that returns "even" if a number in the vector you created above is even and "odd" otherwise.
```{r}
##setup the output
odds_evens <- vector("character", length = length(x))

for (i in seq_along(x)) {
    
    odds_evens[[i]] <- if (x[[i]] %% 2 == 0) { "even" } else { "odd" }
}

odds_evens
```

8. Now complete the code below to do the same thing with an `if_else()` statement and store the results in a variable called odds_evens2

```{r}
odds_evens2 <- if_else(x %% 2 == 0, "even", "odd")

odds_evens2
```

9. Use the `identical()` function to check if the odds_evens is identical to odds_evens2.


```{r}
identical(odds_evens, odds_evens2)
```

```{r}
x + 1
```


```{r}
sum(x)
```


### Task Set 2: Make a for loop

1. Take the gapminder dataset. Loop through its columns. If a column is numeric then store the output of the `summary()` command. If it is not numeric, then store NULL. Print the ouput

2. Use `names(gapminder)` to identify the names of columns and set these as the names of the output vector you created.

### Task Set 3: Functions

1. Create a function that takes two numbers (i.e. atomic vectors of length one) and returns the square of whichever number is the greatest. Use `stop("Your error message here")` to throw an error if the input vector is greater than 1 in length. Test your function to make sure it works.

2. Create a function that takes an atomic vector (of any length) as its input. If it numeric then use the `hist()` function from base to plot a histogram else return an error using `stop()`. See the code below for an illustration of the `hist()` command. Test your function to make sure it works.

```{r}
hist(gapminder$lifeExp)
```



