---
title: "Assignment 1"
author: ~~Hari Subhash~~ Mario B.
date: 'Date created: `r Sys.Date()`'
output:
  html_notebook:
    code_folding: show
    df_print: paged
    highlight: kate
    smart: yes
    theme: cosmo
  html_document: default
---
##Legend
`r emo::ji("check")`: Correct (there might be some improvements I note even with these)

`r emo::ji("man tipping hand")`: Partially correct

`r emo::ji("cross")`: Incorrect answer

`r emo::ji("person gesturing NO")`: Conceptual error that should be avoided.

##Overall Comments

1. Was the assignement submitted on time? `r emo::ji("check")`
2. Was it posted as an online link? `r emo::ji("check")`
3. Does the code follow [recommended style conventions](http://adv-r.had.co.nz/Style.html)? `r emo::ji("check")`
4. Does the notebook compile completely? `r emo::ji("check")`
6. Overall effectiveness of the code? `r emo::ji("check")`: Almost perfect! There are times where you could have written less lines of code for the same result. Look through the solutions for the second part of the assignment.

Scroll to see specific comments.

##Task set 1: Data types and structures
1. List the 4 most common (data) types of atomic vectors.
    + Answer: Logical, integer, numeric or double, and character `r emo::ji("check")`

2. Create atomic vectors of the 4 common types covered in class. Print their length and type. `r emo::ji("check")`

```{r}
log_vec <- c(T)
int_vec <- c(1L)
dbl_vec <- c(1.0)
chr_vec <- c('One')

length(log_vec); typeof(log_vec)
length(int_vec); typeof(int_vec)
length(dbl_vec); typeof(dbl_vec)
length(chr_vec); typeof(chr_vec)
```

3. List the heirarchy of data types for coercion
    + Answer: logical > integer > numeric or double > character `r emo::ji("check")`

4. Explain why the output of following bits of code is TRUE (explain what is happening in terms of coercion): 

```{r}
"1" == 1
```
Answer: 1 is coerced into a character vector `r emo::ji("check")`

```{r}
FALSE == 0
```
Answer: FALSE is coerced into a numeric vector `r emo::ji("check")` 

3. Create a list that holds the 4 common types of atomic vectors you created in task 2. `r emo::ji("check")`
vct_list list()
```{r}
vec_list <- list(log_vec,int_vec,dbl_vec,chr_vec)
vec_list
```
4. Create a data frame with 3 columns and 5 rows using the `data.frame()` function `r emo::ji("check")`

```{r}
df <- data.frame(a = 1:5, b=2, c=3)
df
```
5. Create the same data frame as above using `tibble()` function `r emo::ji("check")`

```{r}
library(tidyverse)
df_tibble <- tibble(a = 1:5, b=2, c=3)
df_tibble
```

6. Use the `str()`, `class()` and or `glimpse()` command to see the difference between the dataframe and tibble create in 4 and 5. `r emo::ji("check")`

```{r}
str(df);str(df_tibble)
class(df);class(df_tibble)
glimpse(df);glimpse(df_tibble)
```
data.frame creates a data.frame whereas tibble creates a tbl_df

7. Fill in the blanks. 
    + A data frame is a _TWO-_ dimensionsional data structure in R that consists of __VECTORS__ of __EQUAL__ length `r emo::ji("check")`

8. What is the difference between an atomic vector and a list?
    + Answer: An atomic vector contains only one data type, but a list can contain multiple types `r emo::ji("check")`

##Task set 2: Data manipulation using dplyr

**INSTRUCTIONS**: You will be using the tidyverse and nycflights13 packages for the next set of tasks. Refer to the documentation on http://tidyverse.org or use the help pages if you do not understand a function fully.

1. Load the tidyverse and nycflights13 packages. `r emo::ji("check")`

```{r}
library(tidyverse); library(nycflights13)
```

2. Describe two ways in which you can use RStudio confirm that a package has been loaded into your current R session? (Hint: One of them is a dropdown menu and the other a list of checkboxes)
    + Answer: In the Environment tab, click the 'Global environment' dropdown and ensure the package is listed or click the Packages and look to ensure the package's box is checked `r emo::ji("check")`

3. Use the `distinct()` command to identify the unique combinations of airline carriers and origin airports in the flights data. `r emo::ji("check")`

```{r}
flights %>%
  distinct(carrier,origin)

```

4. Use the `one_of()` helper function to select the following columns -  year, month, day, dep_time, sched_dep_time, dep_delay, arr_time, from the flights data. `r emo::ji("check")`

```{r}
flights %>%
  select(one_of('year', 'month', 'day', 'dep_time', 'sched_dep_time', 'dep_delay','arr_time'))
```

5. Use the `slice()` function to slice the first 10 rows in the flights data. `r emo::ji("check")`

```{r}
flights %>%
  slice(1:10)
```

6. Perform the same operation as the previous task but this time use `filter()`. (Hint: Search for how to get the row numbers of a tibble) `r emo::ji("check")`

```{r}
filter(flights,row_number() <= 10)
```

7. Use the `top_n()` command to identify the top 5 flights out of NYC that experienced the greatest arrival delays in 2013. `r emo::ji("check")`

```{r}
flights %>%
  top_n(5,arr_delay) %>%
  select(1:3,carrier,flight,arr_delay) %>%
  arrange(desc(arr_delay))
```

8. Use `top_n()` along with `arrange()` and `slice()` to pull out the 50th most delayed departure from NYC in 2013 (Not all 50, just the 50th). `r emo::ji("check")`
```{r}
flights %>%
  top_n(50,dep_delay) %>%
  arrange(desc(dep_delay)) %>%
  slice(50) %>%
  select(1:3,carrier,flight,dep_delay)
  
```

9. Do the same thing as the previous task using 
    + `row_number()` `r emo::ji("cross mark")`: See solution
    + and `dense_rank()` `r emo::ji("cross mark")`
    + <div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: Two issues here - the first is that when you use `flights$dep_delay` you are referring to a version of the `delay` variable that is not piped (within the `dense_rank()` function). Second, you need to use a descending order to rank those with the greatest delays at the top.</div>
```{r}
flights %>%
  filter(dense_rank(flights$dep_delay) == 50) %>%
  select(1:3,carrier,flight,dep_delay)
```

10. The result from `row_number()` and `top_n()` are the same, while that from `dense_rank()` is different. Why?
    + Answer: The dep_delay that falls in the 50th position is tied with several others. In dense_rank, 1749 flights are tied for 50th so it returns all of these. row_number() and top_n() take whatever's 50th `r emo::ji("cross mark")`. See solution.

11. Find the top 10 airline carrier that had the highest average departure delays in 2013 using `group_by()`, `summarise()` and other functions you have learnt previously. `r emo::ji("cross mark")`.

<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: The combo here should be group, summarise, ungroup, top_n. See solution.</div>
```{r}
flights %>%
  group_by(carrier) %>%
  mutate(carr_avg = mean(dep_delay,na.rm = T)) %>%
  select(carrier,carr_avg) %>%
  unique() %>%
  ungroup() %>%
  top_n(10,desc(carr_avg)) %>%
  arrange(desc(carr_avg))
```

12. Use `group_by()` with `mutate()` to create a new variable called `comparativeDelay` which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the `time_hour` variable in the flights data). Store the result in a variable called `comparativeDelays`. `r emo::ji("check")` <div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**:(One small change, you don't need separate mutate commands to create multiple columns, you can do it within a single mutate call.)</div>
```{r}
comparativeDelays <- flights %>%
  group_by(origin,time_hour) %>%
  mutate(avg_delay = mean(dep_delay,na.rm = T)) %>%
  mutate(comparativeDelay = dep_delay - avg_delay) %>%
  ungroup() %>%
  select(origin,carrier,hour,avg_delay,comparativeDelay)

comparativeDelays

```

13. Now group the `comparativeDelays` tibble by carriers to print the top 10 airlines with the greatest average comparative delays.  `r emo::ji("cross mark")`

<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: See 11.</div>
```{r}
comparativeDelays %>%
  ungroup() %>%
  group_by(carrier) %>%
  mutate(avg_comp_delay = mean(comparativeDelay,na.rm = T)) %>%
  select(-origin,-hour,-avg_delay,-comparativeDelay) %>%
  ungroup() %>%
  unique() %>%
  top_n(10,avg_comp_delay) %>%
  arrange(desc(avg_comp_delay))

```

14. Use `group_by()` with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows. `r emo::ji("check")`
```{r}
flights %>%
  group_by(origin) %>%
  top_n(5,dep_delay) %>%
  select(origin,month,day,carrier,flight,dep_delay) %>%
  arrange(origin,desc(dep_delay))
```

15. The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why?
    + Answer: Average comparative delay is best because it's the only one that tries to correct for origin and time-of-day effects`r emo::ji("check")`
    
##Task set 3: R markdown

1. Modify the text below as per the prompt `r emo::ji("check")`
    + Underline me <!--I don't believe I can--> (See solution)
    + **Make me bold**
    + *Make me italic*
    + ~~Strike through me~~

> Make me a block quote

2. Create a level three header called "My level 3 header" under this line `r emo::ji("check")`

### My level 3 header

3. Modify the content below so that the bullet points (unordered list) render correctly`r emo::ji("check")`

**Messy bullet points**

* one
* two
* three

4. Use chunk options to hide the output and the code from the chunk below in the previewed notebook.`r emo::ji("check")`

```{r, results='hide', echo=FALSE}
print("Hide me and the code that created me")
```
