---
title: "Assignment 1"
author: VinPhil
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
##Legend
`r emo::ji("check")`: Correct (there might be some improvements I note even with these)

`r emo::ji("man tipping hand")`: Partially correct

`r emo::ji("cross mark")`: Incorrect answer

`r emo::ji("person gesturing NO")`: Conceptual error that should be avoided.

##Overall Comments

1. Was the assignement submitted on time? `r emo::ji("check")`
2. Was it posted as an online link? `r emo::ji("check")`
3. Does the code follow [recommended style conventions](http://adv-r.had.co.nz/Style.html)? `r emo::ji("man tipping hand")`. Use snake_case instead of titleCase for naming variables.
4. Does the notebook compile completely? `r emo::ji("check")`
6. Overall effectiveness of the code? `r emo::ji("check")`: Almost perfect! There were a few errors but you are almost there.

##Task set 1: Data types and structures
1. List the 4 most common (data) types of atomic vectors. `r emo::ji("check")`
    + Answer: CHar, Int, Double, Logical

2. Create atomic vectors of the 4 common types covered in class. Print their length and type. `r emo::ji("check")`

```{r}
vChar <- c("Hello", "Test")
vInt <- c(1L, 3L)
vDoub <- c(14, 15)
vLog <- c(T,F)

typeof(vChar)
typeof(vInt)
typeof(vDoub)
typeof(vLog)

length(vChar)
length(vInt)
length(vDoub)
length(vLog)
```

3. List the heirarchy of data types for coercion `r emo::ji("check")`
    + Answer: character, double, integer, logical

4. Explain why the output of following bits of code is TRUE (explain what is happening in terms of coercion): `r emo::ji("check")`

```{r}
"1" == 1
```
Answer: == does a comparison and returns T/F. 
For coercion, 1 is converted to character.

```{r}
FALSE == 0
```
Answer: FALSE is converted to 0

3. Create a list that holds the 4 common types of atomic vectors you created in task 2. `r emo::ji("check")`

```{r}
lVectors <- list(vChar, vDoub, vInt, vLog)
lVectors
```
4. Create a data frame with 3 columns and 5 rows using the `data.frame()` function `r emo::ji("check")`

```{r}
dfTest <- data.frame(x=c(1:5), y=c(11:15), z=c(16:20))
dfTest
```
5. Create the same data frame as above using `tibble()` function `r emo::ji("check")`

```{r}
tbTest <- tibble(x = 1:5, y = 11:15, z = 16:20)
tbTest
```

6. Use the `str()`, `class()` and or `glimpse()` command to see the difference between the dataframe and tibble create in 4 and 5. `r emo::ji("check")`

```{r}
str(dfTest)
str(tbTest)

class(dfTest)
class(tbTest)

glimpse(dfTest)
glimpse(tbTest)
```

7. Fill in the blanks. `r emo::ji("man tipping hand")` See solution

    + A data frame is a _2___ dimensionsional data structure in R that consists of _list___ of _equal__ length

8. What is the difference between an atomic vector and a list? `r emo::ji("check")`
    + Answer: atomic verctor is a 1 dimensional array of a single data type - homogeneous. List is also 1d but can be a combimation of different data types - heterogeneous.

<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: Note that array is another data structure in R, so a vector technically is not a one dimensional array but I get what you mean.</div>

##Task set 2: Data manipulation using dplyr

**INSTRUCTIONS**: You will be using the tidyverse and nycflights13 packages for the next set of tasks. Refer to the documentation on tidyverse.org or use the help pages if you do not understand a function fully.

1. Load the tidyverse and nycflights13 packages. `r emo::ji("check")`

```{r}
library(tidyverse)
library(nycflights13)
```

2. Describe two ways in which you can use RStudio confirm that a package has been loaded into your current R session? (Hint: One of them is a dropdown menu and the other a list of checkboxes) `r emo::ji("man tipping hand")`
    + Answer: 1. Global enviroment dropdown. 2. installed.packages()
    + <div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: Technically correct again, and this is down to poor wording of the question. `installed.packages()` does give the list of installed packages, however, the better way to access this through the RStudio GUI is using the packages tab (usually next to help)</div>

3. Use the `distinct()` command to identify the unique combinations of airline carriers and origin airports in the flights data. `r emo::ji("check")`

```{r}
flights
distinct(flights, carrier, origin ) %>%
  arrange(carrier, origin)
```

4. Use the `one_of()` helper function to select the following columns -  year, month, day, dep_time, sched_dep_time, dep_delay, arr_time, from the flights data. `r emo::ji("check")`

```{r}
flights %>% 
  select(one_of(c("year", "month", "day", "dep_time", "sched_dep_time", "dep_delay", "arr_time")))

```

5. Use the `slice()` function to slice the first 10 rows in the flights data. `r emo::ji("check")`

```{r}
flights %>% 
  slice(1:10)
```

6. Perform the same operation as the previous task but this time use `filter()`. (Hint: Search for how to get the row numbers of a tibble) `r emo::ji("check")` <span style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**:(This works, another solution would be `filter(row_number() <= 10)`)</span>

```{r}
flights %>% 
  filter(between(row_number(), 1, 10))
```

7. Use the `top_n()` command to identify the top 5 flights out of NYC that experienced the greatest arrival delays in 2013. `r emo::ji("check")`
```{r}
flights %>% 
  top_n(5, arr_delay) %>% 
  arrange(desc(arr_delay))
```

8. Use `top_n()` along with `arrange()` and `slice()` to pull out the 50th most delayed departure from NYC in 2013 (Not all 50, just the 50th). `r emo::ji("check")`
```{r}
flights %>% 
  top_n(50, dep_delay) %>% 
  arrange(desc(dep_delay)) %>% 
  slice(50)
```

9. Do the same thing as the previous task using `r emo::ji("check")`
    + `row_number()`
    + <div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: You don't need to use the arrange command here. `row_number()` is a ranking function just like `dense_rank()`. So you can do this `filter(row_number(desc(dep_delay)) == 50)` this would use the `dep_delay` as the variable to generate the row numbers on.</div>
    + and `dense_rank()`
```{r}
flights %>% 
  arrange(desc(dep_delay)) %>% 
  filter(row_number() == 50)
```

```{r}
flights %>% 
  filter(dense_rank(desc(dep_delay)) == 50)
```


10. The result from `row_number()` and `top_n()` are the same, while that from `dense_rank()` is different. Why? `r emo::ji("check")`
    + Answer: dense rank assigns the same rank for the same value of dep_delay, whereas the others are sequentially ranked

11. Find the top 10 airline carrier that had the highest average departure delays in 2013 using `group_by()`, `summarise()` and other functions you have learnt previously. `r emo::ji("check")`

```{r}
flights %>% 
  group_by(carrier) %>% 
  summarise(avgDepDelay = mean(dep_delay, na.rm = T)) %>% 
  top_n(10, avgDepDelay) %>% 
  arrange(desc(avgDepDelay)) 
```

12. Use `group_by()` with `mutate()` to create a new variable called `comparativeDelay` which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the `time_hour` variable in the flights data). Store the result in a variable called `comparativeDelays`. `r emo::ji("cross mark")`

<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: You want to create a dataset that has one row for every carrier for every hour in 2013, this would require a `group_by() %>% summarise()` combo. You also need to use the `time_hour` variable and not the `hour` since hour repeats every day (i.e. it goes from 1 to 24 every day) while `time_hour` is a date object that records every hour. See solution for the correct implementation.</div>
```{r}
flights %>% 
  group_by(origin, hour) %>% 
  mutate(avgDelay = mean(dep_delay, na.rm = T), 
         comparativeDelay = dep_delay - mean(dep_delay, na.rm = T)) %>% 
  select(1:3, carrier, hour, flight, tailnum, dep_delay, avgDelay, comparativeDelay)

```

13. Now group the `comparativeDelays` tibble by carriers to print the top 10 airlines with the greatest average comparative delays. `r emo::ji("cross mark")` (See solution)
```{r}
flights %>% 
  group_by(origin, hour) %>% 
  mutate(avgDelay = mean(dep_delay, na.rm = T), 
         comparativeDelay = dep_delay - mean(dep_delay, na.rm = T)) %>% 
  select(1:3, carrier, hour, flight, tailnum, dep_delay, avgDelay, comparativeDelay) %>% 
  group_by(carrier) %>% 
  summarise(avgComparativeDelay = mean(comparativeDelay, na.rm = T)) %>% 
  top_n(10, avgComparativeDelay) %>% 
  arrange(desc(avgComparativeDelay))
```

14. Use `group_by()` with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows. `r emo::ji("man tipping hand")`
<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: Good job figuring this out. The easier way to do this would be to the following combo: `group_by() %>% top_n()`</div>

```{r}
flights %>%
  group_by(origin) %>% 
  filter(between(dense_rank(desc(dep_delay)),1, 5)) %>% 
  arrange(origin, desc(dep_delay))

```

15. The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why? `r emo::ji("check")`
    + Answer: Average Comparative Delay seems to be the best, as it summarises by origin and hour 
    
##Task set 3: R markdown

1. Modify the text below as per the prompt `r emo::ji("check")`
    + Underline me
    + **Make me bold**
    + >Make me a block quote
    + *Make me italic*
    + ~~Strike through me~~

2. Create a level three header called "My level 3 header" under this line `r emo::ji("cross mark")`
###My level 3 header

<div style = "background-color:rgba(244, 155, 53, 0.5)">**Hari**: You need to add a new line before the heading (since the content above it is a list item.</div>

3. Modify the content below so that the bullet points (unordered list) render correctly `r emo::ji("check")`

**Messy bullet points**

- one
- two
- three

4. Use chunk options to hide the output and the code from the chunk below in the previewed notebook. `r emo::ji("check")`

```{r, results='hide', echo=FALSE}
print("Hide me and the code that created me")
```