---
title: "Class 22 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---
## Functions
The code below creates a function that takes two numeric vectors (atomic) as its input and returns the absolute difference between the minimum value of the first vector and the maximum value of second vector. Read it carefully (along with the comments).

```{r}
minmax_diff <- function(min_vec, max_vec, na.rm = F) {
  ##add a stop condition
  # if (!is.numeric(min_vec)) {
  #   stop("Hari's error message: min_vec should be numeric")
  # }
  
  stopifnot(is.numeric(min_vec), is.numeric(max_vec), is.logical(na.rm))
  
  # if (!is.logical(na.rm)) {
  #   na.rm <- T
  # }
  
  ##Lets calculate the min and max values
  min_val <- min(min_vec, na.rm = na.rm)
  max_val <- max(max_vec, na.rm = na.rm)
  
  ## Let calculate the value that needs to be returned
  abs(min_val - max_val)
  
}
minmax_diff(1, 2, F, 6)
```

```{r}
minmax_diff(1:23, c(4, NA), na.rm = "xyz")
```


```{r}
mean(c(4, NA), na.rm = "xyz")
```

```{r}
as.logical("xyz")
```

```{r}
if("xyz") {
  print("Truthy")
}
```


```{r}
min(min_vec, na.rm = na.rm)
```


```{r}
minmax_diff(min_vec = 1:23, max_vec = c(4, NA), na.rm = T)
```


```{r}
minmax_diff(min_vec = c(24, 41, NA), max_vec = c(44, NA), na.rm = T)
```


```{r}
minmax_diff(min_vec = "a", max_vec = c(44, NA), na.rm = T)
```


```{r}
min(c(24, 41, NA))
```

```{r}
minmax_diff(1:10, 20:30)
```


Let's test to see if it works.
```{r}
minmax_diff(min_vec = 1:10, max_vec = 10:20)
```

Introducing NA values.
```{r}
minmax_diff(min_vec = c(24, 41, NA), max_vec = 44)
```

Specifiying the NA condition.
```{r}
minmax_diff(min_vec = c(24, 41, NA), max_vec = 44, na.rm = T)
```

Trying out a non-numeric input.

```{r}
minmax_diff(min_vec = "a", max_vec = c(1, 2, 56))
```

1. How does this code work?
```{r}
minmax_diff(1:23, 4, T)
```

2. Why does this give an error? 

```{r}
minmax_diff(1, 2, 5)
```

3. Why does this not work?

```{r}
minmax_diff(1, 2, F, 6)
```

4. Define a function that takes two numeric vectors as its input and returns the difference between the mean of the first vector and the median of the second vector. Add any checks and additional arguments you like to this function.
```{r}
diff_vec <- function(mean_vec, median_vec, na.rm){
  ## stop condition
  stopifnot(is.numeric(mean_vec), is.numeric(median_vec), is.logical(na.rm))
  
  ## calculating the value to be returned
  mean(mean_vec, na.rm = na.rm) - median(median_vec, na.rm = na.rm)
}

diff_vec(mean_vec = 100:200, median_vec = 20:30, na.rm = T)
```

5. Define a function that takes two arguments - name and greet_message. Set the default value of greet_message to be "Hello". The function should return a single character string that combines greet_message with name (with a space in between). Use `paste()` to combine. Add any additional arguments or checks.

```{r}
# greet_message <- "Hi"
greet_person <- function(name, greet_message = greet_message) {
  paste(greet_message, name)
}
# TODO debug below
# greet_person(name = "Bobby")

greet_person(name = "Bobby", greet_message = "How are you")
```

```{r}
greet_person(name = "Bobby", greet_message = greet_message)
```

```{r}

greet_person(name = "Bobby", greet_message = greet_message)
```


This will not work. Why?
```{r}
greet_person(Hari)
```

This will. Why?
```{r}
Bobby <- "Hari"
greet_person(name = Bobby)
```

6. Load the tidyverse and gapminder packages
```{r}
library(tidyverse); library(gapminder)
```

7. Fix the following so that they run. Think about why they are generating an error. Read the comments for further instructions.



```{r}
##Find the unique set of continents in gapminder
##Your final solution should use pipes
gapminder %>% 
  unique(continent)
```
The same error in base R. 
```{r}
unique(gapminder, continent)
```

```{r}
gapminder %>% 
  select(continent) %>% 
  unique()
```

if you want an atomic vector as your output
```{r}
gapminder %>% 
  pull(continent) %>% 
  unique()
```


without pipes
```{r}
unique(gapminder$continent)
```


```{r}
##find the mean of the gdpPercap column in gapminder
##final solution should use pipes.
gapminder %>% 
  select(gdpPercap) %>% 
  mean(., na.rm = T)
```

```{r}
gapminder %>% 
  pull(gdpPercap) %>% 
  mean(., na.rm = T)
```

```{r}
## find the mean of the gdpPercap column in gapminder
## final solution should use base R syntax 
## remember [] - is a preserving subset | $ and [[]] are simplifying
mean(gapminder["gdpPercap"])
```

```{r}
mean(gapminder[["gdpPercap"]])
```

```{r}
## I would like the distinct combinations of continent and country
## use tidyverse
gapminder %>% 
  distinct(continent, country)
```

```{r}
## I would like to have a single column tibble that contains all the distinct continents
## Use tidyverse
gapminder %>% 
  pull(continent) %>% 
  distinct()
```

