---
title: "Class 22 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---
## Functions
The code below creates a function that takes two numeric vectors as its input and returns the absolute difference between the minimum value of the first vector and the maximum value of second vector. Read it carefully (along with the comments).

```{r}
##Defining the function (notice how I also define the na.rm.)
minmax_diff <- function(min_vec, max_vec, na.rm = F) {
  ##check if the two arguments are of the type numeric, else exit.
  stopifnot(is.numeric(min_vec), is.numeric(max_vec), is.logical(na.rm))
  
  ##get the min and max value.
  min_value <- min(min_vec, na.rm = na.rm)
  max_value <- max(max_vec, na.rm = na.rm)
  
  ##return the absolute difference
  abs(max_value - min_value)
  
}
```

Let's test to see if it works.
```{r}
minmax_diff(min_vec = 1:10, max_vec = 10:20)
```

Introducing NA values.
```{r}
minmax_diff(min_vec = c(24, 41, NA), max_vec = 44)
```

Specifiying the NA condition.
```{r}
minmax_diff(min_vec = c(24, 41, NA), max_vec = 44, na.rm = T)
```

Trying out a non-numeric input.

```{r}
minmax_diff(min_vec = "a", max_vec = c(1, 2, 56))
```

1. How does this code work?
```{r}
minmax_diff(1:23, 4, T)
```

2. Why does this give an error? 

```{r}
minmax_diff(1, 2, 5)
```

3. Why does this not work?

```{r}
minmax_diff(1, 2, F, 6)
```

4. Define a function that takes two numeric vectors as its input and returns the difference between the mean of the first vector and the median of the second vector. Add any checks and additional arguments you like to this function.
5. Define a function that takes two arguments - name and greet_message. Set the default value of greet_message to be "Hello". The function should return a single character string that combines greet_message with name (with a space in between). Use `paste()` to combine. Add any additional arguments or checks.
6. Load the tidyverse and gapminder packages
7. Fix the following so that they run. Think about why they are generating an error. Read the comments for further instructions.
```{r}
##Find the unique set of continents in gapminder
##Your final solution should use pipes
gapminder %>% 
  unique(continent)
```
```{r}
##find the mean of the gdpPercap column in gapminder
##final solution should use pipes.
gapminder %>% 
  select(gdpPercap) %>% 
  mean(., na.rm = T)
```


```{r}
## find the mean of the gdpPercap column in gapminder
## final solution should use base R syntax 
## remember [] - is a preserving subset | $ and [[]] are simplifying
mean(gapminder["gdpPercap"])
```

```{r}
## I would like the distinct combinations of continent and country
## use tidyverse
gapminder %>% 
  unique(continent, country)
```

```{r}
## I would like to have a single column tibble that contains all the distinct continents
## Use tidyverse
gapminder %>% 
  pull(continent) %>% 
  distinct()
```


## Group and do things
1. Use summarise to calculate the mean population for entire gapminder dataset. Compare this value with `mean(gapminder$pop)` (you will have to run this to see the output).
2. Repeat the same step after grouping by continent.
3. Repeat the same after grouping by continent and year.
4. Use `n_distinct()` to find the number of distinct countries in the gapminder data. 
5. Use any of the simplifying subset operators (`$`, `[[]]` or `pull()`) to pull the countries variable from gapminder. See the differences with 4.
6. Combine 3 and 4 to create a table that shows the distinct number of countries and the mean population in every continent for each year. You can create multiple summarised columns by separating them within the summarise command using commas 
`summarise(var1 = some code, var2 = some code)`.
6. Now group the data to show the distinct number of countries in every continent for every year. Use `n_distinct()` to find the number of distinct countries/continents.
7. Group the gapminder dataset and filter the countries that have the highest or lowest lifeExp for each continent for each year. Arrange the output by year,  continent and lifeExp.
8. Filter the most recent year of the data and remove the continent Oceania from it. Create a new variable that is marked "Above average income" if the gdpPercap is greater than the average gdpPercap for the continent it belongs to and "Below average income" if it is below. Use `group_by()` with `mutate()` for this.


## ggplot

1. The code below creates a scatter plot (geom_point()). Update the aesthetics so that the colour of the dots reflect the continents.

```{r}
ggplot(gapminder, aes(x = year, y = pop)) +
    geom_point()
```

2. Update this plot by adding a title using `labs()`. Use the `title` argument to set the title.
3. The code below removes the x-axis title "year" by setting the element to blank (see the difference from above). Update this by also removing the y-axis label using `axis.title.y`.

```{r}
ggplot(gapminder, aes(x = year, y = pop)) +
    geom_point() +
    theme(axis.title.x = element_blank())
```

4. The code below changes the angle of the x axis labels. Replicate this for the y-axis. Set whatever you would like. 

```{r}
ggplot(gapminder, aes(x = year, y = pop)) +
    geom_point() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45))
```

5. Go to the tidyverse reference for ggplot to see other parameters that you can control within element_text. Use this to also set the color of the y-axis labels.
6. In the code below I use an external variable to set the colour based on whether a country is India/China or not. Use the same logic to set the alpha parameter to either 1 if India/China or 0.3 otherwise.

```{r}
india_china_color <- if_else(gapminder$country %in% c("China", "India"), "pink", "black")
ggplot(gapminder, aes(x = year, y = pop)) +
    geom_point(colour = india_china_color) 
```





