---
title: "Bobby"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---

```{r}
library(tidyverse); library(nycflights13)

skimr::skim(flights)
```

```{r}
skimr::skim(weather)
```


##Effect of temp on delay

```{r}
ggplot(weather, aes(x = temp)) +
    geom_histogram()
```

Lets see how many distinct values there are in the temperature column
```{r}
n_distinct(weather$temp)
```

So there don't seem to be too many unique values in temp. Lets print those out to just see them.
```{r}
unique(weather$temp)
```

lets aqlso look at the number of missing values in temp
```{r}
sum(is.na(weather$temp))
```

Since there are only 174 unique values of temp it is not unreasonable to group based on temp.
```{r}
flights %>% 
    filter(dep_delay > 20) %>%
    select(origin, time_hour, dep_delay) %>% 
    left_join(weather, by = c("origin", "time_hour")) %>% 
    group_by(temp) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n()) %>% 
    ggplot(aes(x = temp, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm")
```

Lets now mark the low observation values.
```{r}
flights %>% 
    filter(dep_delay > 20) %>%
    select(origin, time_hour, dep_delay) %>% 
    left_join(weather, by = c("origin", "time_hour")) %>% 
    group_by(temp) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n(),
              smallSample = if_else(n() < 15, "small sample", "large sample")) %>% 
    ggplot(aes(x = temp, y = meanDelay)) +
    geom_point(aes(colour = smallSample)) +
    geom_smooth(method = "lm")
```

Some of the variation seems to be driven by small samples. So lets remove those.
```{r}
flights %>% 
    filter(dep_delay > 20) %>%
    select(origin, time_hour, dep_delay) %>% 
    left_join(weather, by = c("origin", "time_hour")) %>% 
    group_by(temp) %>% 
    #only keep groups with more than 15 obs.
    filter(n() > 15) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = temp, y = meanDelay)) +
    geom_point(aes()) +
    geom_smooth(method = "lm")
```

Lets show you how to cut.

```{r}
flights %>% 
    filter(dep_delay > 20) %>%
    select(origin, time_hour, dep_delay) %>% 
    left_join(weather, by = c("origin", "time_hour")) %>% 
    mutate(tempGroup = cut_number(temp, n = 10)) %>% 
    ggplot(aes(x = tempGroup, y = dep_delay)) +
    geom_jitter(alpha = 0.1, aes(colour = origin)) +
    geom_boxplot(outlier.shape = NA) +
    coord_flip() +
    scale_y_continuous(limits = c(0, 200))
```





