---
title: "Make up class 1 - Exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
**INSTRUCTIONS**:

1. Give your answers below each numbered task. Enter a new line.
2. The tasks are clustered under level 2 headers (`##`). Do one set at a time.
3. Preview your notebook at the end of each in class exercise task set. Does the notebook look formatted correctly? If not fix it.

**KEYBOARD SHORTCUTS**:

1) Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2) Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3) Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4) Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5) Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

##Task set 1: Create your folder structure
1. Create the folder structure for the class on your computer called "yourfirstnameR4D" with the following subfolders in it - "Assignments", "Classes" and "Data". This will be your folder for the entire course.
2. Download and save the Rmd file using the drop down at the top right of this notebook in a subfolder called "Makeup Class1" in the "Classes" folder.
3. Create your first code chunk (remember to use the shortcuts) under this line
5. Within it assign any value you would like to a variable. Print this value. Save the file in the "Classes" subfolder
6. Click on the preview button to see the notebook. 
7. Add a level two heading using `##` under this line with a short two sentence paragraph under it explaining what R markdown is.
8. Change the author in the YAML section to your name.
9. Hit preview again to see the differences from the first time.

##Task set 2: Data structures and types
1. Create and store a character vector (of your choice), a numeric vector (interger/double) and a factor, all of the same length
2. Get the type (using `typeof()`), and the length (using `length()`) of the vectors you created in 1.
3. Create a list containing all the vectors you created in 1, using `list()`.
4. Try combining using `c()` instead of `list()`. Can you explain why the output from the two commands are different?
5. Create a dataframe using all the vectors stored in step 1
6. Predict the output of the following without running the code in 
    + `c("FALSE", FALSE, T, F, T, T, T)` = "Character", 
    + `c(1.0, FALSE)` = "numeric", 
    + `c(2, TRUE)`  = "numeric"and 
    + `c("two", 2)` = "character"
    
##Task set 3: Package, arrange, select and mutate
1. Install and load the `nycflights13` package and the `tidyverse` package
2. Arrange the flights data from the NY in descending order of dates (the last day of the year should be the first row of the data).
3. Use the documentation for `select()` on the tidyverse portal to find a way to create a new tibble that contains the first three columns (year, month, day) and all other columns whose names `contains()` the term "delay"
4. Create a column that indicates whether a flight took off in the AM or the PM, using `if_else()`. Make sure to have the help file for `if_else()` open when you do this.

##Task set 4: Filter
1. Filter the flights data to only show the flights in the month of December
```{r}
flights %>% 
    filter(month == 12)
```

2. Filter all the flights with departure that were delayed by more than 2 hours in the month of December
```{r}
flights %>% 
    filter(month == 12 & dep_delay > 120)
```

3. Show all the rows for which `arr_delay` is missing (`NA`).
```{r}
flights %>% 
    filter(is.na(arr_delay))
```

4. Filter the flights that were between the 10th and 40th most delayed in terms of arrival (`arr_delay`) using the `dense_rank()` helper function.
```{r}
flights %>% 
    mutate(arrDelayRank = dense_rank(desc(arr_delay))) %>% 
    select(1:3, arr_delay, arrDelayRank) %>% 
    arrange(desc(arr_delay)) %>% 
    filter(arrDelayRank %in% 10:40)
```

##Task set 5: Grouping
1. Use `group_by()` and `summarise()` to find the average arrival delay to different destinations (`dest`).
2. Create a new variable in the flights data that is difference between the departure delay for each flight and the monthly average delay.
3. Use `group_by()` and `filter()` to identify the flights that had the highest delay for every month in 2013.





