---
title: "Make up class 1 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
##Learning objectives
1. ~~Master the basics of variable assignment~~
2. ~~Familiarize with RStudio, notebooks, markdown and YAML~~
3. ~~Understand data structures and types in R~~
4. ~~Create atomic vectors, identify data types and rules of coercion~~
5. ~~Create lists and data frames~~
6. ~~Understand the package system in R~~
7. ~~Install and load packages in R~~
8. ~~Comprehend the basics of the tidyverse (the pipe, tibble, basic principles)~~
9. ~~Arrange a tibble based on column values~~
10. ~~Select columns from a tibble~~
11. ~~Create new columns in a tibble~~
12. ~~Filter rows from a data frame~~
13. Master grouping
14. Create an online notebook via GitHub


##Some R markdown practice

##Here is a test heading
**bold text**

*italic*

Unordered list (remember to enter a new empty line if the line right above an list is not a heading)

* list item 1
* list item 2
* list item 3

ordered list

1. item 1
2. item 2
3. item 3
    + subitem 1
    + subitem 2
    
YAML: Yet Another Markup Language
    
##Variable Assignment
Lets assign a value to our first variable. 
```{r}
myFirstVar <- "Hello World!"
```

The name of the variable is on the left side of the assignment operator. The assignment operator is created using the shortcut `CMD + Option + I` (Ctrl + Alt + I for Windows)

Nows let print/call the variable that we created
```{r}
myFirstVar
```

Let reassign a value to `myFirstVar`.
```{r}
myFirstVar <- 23
```

Create a new variable and assign it the value from the first.
```{r}
mySecondVar <- myFirstVar
```


##Create atomic vectors
Lets create atomic vectors of different data types.
```{r}
chrVec <- c("Mario", "Adiam", "Hari")
numVec <- c(1, 2, 4, 6)
logicalVec <- c(T, F, TRUE, F)
```

Vectors can be created easily using the `c()` function which is a shortform for "combine".

Lets get the type and length of the vectors we created.
```{r}
typeof(chrVec)
typeof(numVec)
typeof(logicalVec)
```

```{r}
intVec <- c(1L, 3L, 6L, 9L)
typeof(intVec)
```
Now lets get the lengths.
```{r}
length(chrVec)
```


###use a few convenience functions to create vectors

`seq()`
```{r}
sampleSeq <- seq(from = 10, to = 30, by = 2)
sampleSeq
```

`:`
```{r}
10:30
```


###Coercion
The heirarchy of coercion in R is as follows: character > numeric > integer > logical
```{r}
c("one", 1, 1.2)
```

```{r}
dblInt <- c(1.2, 3L, 3.4)
typeof(dblInt)
```

```{r}
is.numeric(dblInt)
```

```{r}
logicalNum <- c(T, 1, F)

typeof(logicalNum)
```

##Install and load packages

```{r}

```

##Basics of the tidyverse

Pipe operator `%>%`.

Tibbles are data.frames ++

```{r}
library(nycflights13); library(tidyverse)

flights
```

```{r}
str(flights)
```


```{r}
mtcars
```

```{r}
str(mtcars)
```

Its easy to convert a data.frame to a tibble.

```{r}
tblMtCars <- as_tibble(mtcars)
```


```{r}
str(tblMtCars)
```

##Arrange a tibble
You can arrange a tibble using the `arrange()` verb.

```{r}
flights %>% 
    arrange(desc(month))
```

You can save this arranged tibble to an object using the <- assignment operator.
```{r}
sampleSave <- flights %>% 
    arrange(desc(month))
```

If you would like to create a new copy of flights in your global environment that is arranged, you can do that as well
```{r}
flights <- flights %>% 
    arrange(desc(month))
```

The code below removes flights from your environment.
```{r}
rm(flights)
```

##Select columns
Select using column positions
```{r}
flights %>% 
    select(1, 2, 4)
```

You can also use the names of the columns to select them

```{r}
flights %>% 
    select(year, month, dep_delay)
```

I can also use a range using names of columns
```{r}
flights %>% 
    select(year:dep_delay)
```
Lets try selecting using a helper function `starts_with()`
```{r}
flights %>% 
    select(starts_with("arr"))
```


##Mutate to add new columns
You can add new columns to a tibble using `mutate()`
```{r}
flights %>% 
    select(year, month, day) %>% 
    mutate(newDate = paste(year, month, day, sep = "-"))
```

Lets use an `if_else()` within a mutate
```{r}
flights %>% 
    select(1:3, dep_delay) %>% 
    mutate(delayedOrNot = if_else(dep_delay > 0, 
                                  "Flight departure was delayed", 
                                  "Flight departed on time or early"))
```




```{r}
flights
```

##Filtering rows from a tibble

Lets test out a simple case.
```{r}
flights %>% 
    filter(month == 12)
```

1. Equalization operator `==`

```{r}
2 == 2
3 == 4
```

```{r}
3 > 4
```

```{r}
3 < 4
```

```{r}
2 != 2
```

```{r}
4 < 4
4 <= 4
```

```{r}
numVect <- 20:60
numVect < 30
```

Filtering using multiple conditions

```{r}
flights %>% 
    filter(month == 12 & distance > 1500) %>% 
    select(1:3, distance)
```

Lets try using an or condition to filter rows
```{r}
flights %>% 
    filter(month == 12 | month == 1) %>% 
    head()
```

Lets find 10th longest flight by distance in the year 2013
```{r}
flights %>% 
    filter(dense_rank(desc(distance)) == 10) %>% 
    select(1:3, distance)
```

I want pull out the flights that are ranked 10th to 20th in terms of distance

```{r}
flights %>% 
    mutate(distanceRank = dense_rank(desc(distance))) %>% 
    filter(distanceRank %in% 10:20) %>% 
    select(1:3, distance, distanceRank)
```



```{r}
namesOfParticipants <- c("Diana", "Mario", "Erika", "Marc", "Adiam")

namesOfParticipants %in% c("Mario", "Marc")
```


```{r}
numVect %in% 20:35
```


```{r}
numVect < 36
```

##Grouping
Lets look at a simple grouping
```{r}
flights %>% 
    group_by(origin) %>% 
    summarise(avgFlightDistance = mean(distance, na.rm = T))
```
lets try grouping with two groups
```{r}
flights %>% 
    group_by(month, origin) %>% 
    summarise(avgFlightDistance = mean(distance, na.rm = T))
```

mutate, and filter with grouping
```{r}
flights %>% 
    group_by(carrier) %>% 
    mutate(meanDelay = mean(dep_delay, na.rm = T), 
           comparativeDelay = dep_delay - meanDelay) %>% 
    select(carrier, dep_delay, meanDelay, comparativeDelay)
```

filter with grouping
```{r}
flights %>% 
    group_by(origin) %>% 
    mutate(maxDelay = max(dep_delay, na.rm = T)) %>% 
    filter(dep_delay == maxDelay) %>% 
    select(1:3, origin, maxDelay, dep_delay)
```

