---
title: "Class 18 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

### Task Set 1: For loop

Go here to check out the [basic elements of a for loop](https://puzhu.github.io/R4D/Classes/Class%2017/class17_slides.html#/a-common-recipe). 

**TASK 1**: Load the gapminder dataset. Loop through its columns. If a column is numeric then store the output of the `summary()` command. If it is not numeric, then store NULL. Print the output.

```{r}
##make your loop here
```


You can use `names()` to both get and set the names of a vector. The code below `gets` the names of the `mtcars` dataset.

```{r}
names(mtcars)
```

In the code below I use `names()` to set the column headers of a copy of the `mtcars` dataset called `mtcars_2`. See the output to notice how the column headings have been changed.
```{r}
## create a copy of the mtcars data
mtcars_2 <- mtcars

names(mtcars_2) <- paste0(names(mtcars), "_2")

mtcars_2
```

**TASK 2**: Replicate the code to get `names()` of gapminder and append the string "_out" to each of them and set them as the names of the output vector you created in task 1.

### Task Set 2: Functions

**Task 1**: Create a function called `square_greater()` that takes two arguments x and y that are numbers (i.e. atomic vectors of length one) and returns the square of whichever number is the greatest. Use `stopifnot(condition1, condition2)` to throw an error if any of the two input values are greater than 1 in length. 
```{r}
##define your function here
square_greater <- function(){
    
}
```


Test your function to make sure it works.
```{r}
##run this after you define your function
square_greater(x = 32, y = 87) ## this should work

square_greater(x = 3.2, y = "4") ## this should work

square_greater(x = 23:45, y = 89) ## this should throw an error
```

The `hist()` command is from base R and takes a numeric vector as its input. See the code below for an illustration using the `wt` column from the `mtcars` dataset.
```{r}
hist(mtcars$wt)
```

**Task 2**: Create a function called `draw_hist()` that takes an atomic vector (of any length) as its input. If it numeric then use the `hist()` function from base to draw a histogram else return an error with a custom message of your choosing using `stop()`. 

```{r}
##create your function
draw_hist
```


Test your function to make sure it works.
```{r}
library(gapminder)
latest_gap <- gapminder %>% 
                   filter(year == max(year, na.rm = T))

draw_hist(pull(latest_gap, lifeExp)) ## this should plot a histogram of life expectancy
draw_hist(rnorm(n = 1000, mean = 20, sd = 4)) ## this should plot a histogram of a numeric vector with mean 20 and std dev of 4
draw_hist(pull(latest_gap, continent)) ## this should throw an error
```





