---
title: "Class 9 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
**LEARNING OBJECTIVE**: Combine different relational tables using different types of data joins.

**TOPICS TO COVER**

1. What is relational data?
2. Types of keys
3. Types of joins
    + Mutating joins
    + Filtering joins
4. Joining by multiple keys with different names using a named character vector
5. Handling duplicate keys

## Why use relational data?

Relational data maintains data integrity and saves space. It also makes maintenance of datasets much more easier.

All the datasets in the flights data are linked to each other through their keys. See prep material for more information on this. I have printed out a few of the datasets from `nycflights13` to illustrate.
```{r}
library(nycflights13); library(tidyverse)
flights
```

```{r}
airlines
```

```{r}
airports
```

##Mutating joins
Lets illustrate a left_join
```{r}
flights %>% 
    select(carrier, dep_delay) %>% 
    left_join(x = ., y = airlines, by = "carrier")
```

In case of duplicates we get more rows than before. See how there is an additional row for UA now that corresponds to the row we added.
```{r}
##I have used bind_rows to add a row to the airlines data ()
duplicatedAirlines <- bind_rows(airlines, 
                                tibble(carrier = "UA", 
                                       name = "The new United Airline")
                                )

tail(duplicatedAirlines)
```
Lets see what happens when we join this new duplicated data.
```{r}
flights %>% 
    select(carrier, dep_delay) %>% 
    left_join(x = ., y = duplicatedAirlines, by = "carrier")
```

Now we have more rows that the original flights data because we have duplicated the rows where there are matches. Every single row of UA is now duplicated.

What happens when your keys are not called the same names?

Lets rename the carrier as newCarrier.
```{r}
notSameKey <- airlines %>% 
    ##we can use select to rename  columns as well
    select(newCarrier = carrier, name)

notSameKey
```

Lets say we now want to do the same left join as before with this new data
```{r}
flights %>% 
    select(carrier, dep_delay) %>% 
    left_join(x = ., y = notSameKey, by = c("carrier" = "newCarrier"))
```



