---
title: "Assignment 1"
author: "Hari Subhash"
date: 'Date created: `r Sys.Date()`'
output:
  html_notebook:
    code_folding: show
    df_print: paged
    highlight: kate
    smart: yes
    theme: cosmo
  html_document:
    df_print: paged
---

**Due Date**: Sunday, March 24th

**Keyboard Shortcuts**:

1. Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2. Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3. Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4. Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5. Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

**General Instructions**

1. Use the cheatsheets and the keyboard shortcuts as you work through the assignment.
2. Each task should be answered right below the question.
3. The assignment should be submitted via a shareable github link on the assignments channel in Slack.
4. Make sure the notebook is neatly formatted and looks good when you preview it.

##Task set 1: Data types and structures
1. List the 4 most common (data) types of atomic vectors.
    + Answer: *Logical, Integer, Double, Character*.

2. Create atomic vectors of the 4 common types covered in class. Print their length and type.

```{r}
AV1 <- c("Chip", "Widget", "girdle", "burgle", "fork")
AV2 <- c(43L, 32L, 24L, 16L)
AV3 <- c(33.5, 14.7, 23.434334, 43/79)
AV4 <- c(T, F, F, T, F)
typeof(AV1)
typeof(AV2)
typeof(AV3)
typeof(AV4)
length(AV1)
length(AV2)
length(AV3)
length(AV4)
```

3. List the heirarchy of data types for coercion
    + Answer: *Logical, integer, double, character*. To explain further, in a vector, types to the left will be converted into their rightward types, if mixed types exist.

4. Explain why the output of following bits of code is TRUE (explain what is happening in terms of coercion): 

```{r}
"1" == 1
"D" == "D"
```
Answer: I believe that the integer value is being coerced into a character value, thus leaving identical values on each side, thus the equality 
evaluates to true.

```{r}
FALSE == 0
```
Answer: There are mixed vector data types here. As per the coercion hierarchy, the logical value "FALSE" is coerced to match the (integer?) type of the right side of the equality. Thus FALSE is coerced to 0, and the equality evaluates to TRUE.

3. Create a list that holds the 4 common types of atomic vectors you created in task 2.

```{r}
Hold4List <- list(AV1, AV2, AV3, AV4)
```
4. Create a data frame with 3 columns and 5 rows using the `data.frame()` function

```{r}
SignDF001 <- data.frame(num = 1:5, char = AV1, logical = AV4 )

SignDF001
#note listed in environment as 5 "observations" (rows) of 3 "variables" (columns)
```
5. Create the same data frame as above using `tibble()` function

```{r message = FALSE}
library(tidyverse)
SignTIB01 <- tibble(num = 1:5, char = AV1, logical = AV4)
```

6. Use the `str()`, `class()` and or `glimpse()` command to see the difference between the dataframe and tibble create in 4 and 5.
  ANSWER: Tibble treats char vectors as such; dataframe converts them to factors. tibl is 3 classes, df is one.
```{r}
str(SignTIB01)
str(SignDF001)
#BREAK
class(SignTIB01)
class(SignDF001)
#BREAK
glimpse(SignTIB01)
glimpse(SignDF001)

```

7. Fill in the blanks. 

    + A data frame is a __two?__ dimensionsional data structure in R that consists of ___columns_ of _equal___ length

8. What is the difference between an atomic vector and a list?
    + Answer: Many differences? A vector has direction, i.e. the elements have a defined order, where a list's elements do not. A vector must have homogenous data types, but a list need not. Lists can be recursive and contain other lists, while vectors can't. 


##Task set 2: Data manipulation using dplyr

**INSTRUCTIONS**: You will be using the tidyverse and nycflights13 packages for the next set of tasks. Refer to the documentation on tidyverse.org or use the help pages if you do not understand a function fully.

1. Load the tidyverse and nycflights13 packages.

```{r}
#install.packages("tidyverse")
#install.packages("nycflights13")
library(tidyverse)
library(nycflights13)
```

2. Describe two ways in which you can use RStudio confirm that a package has been loaded into your current R session? (Hint: One of them is a dropdown menu and the other a list of checkboxes)

    + Answer:
    the first method is the "global environment" sub-tab under the environment tab. Click on it and look for your pacakge.
    the second method is the "Packages" tab, which shows a list of packages in both user and system libraries.

3. Use the `distinct()` command to identify the unique combinations of airline carriers and origin airports in the flights data.

```{r}
flights %>% 
distinct(carrier, origin) %>% 
  arrange(carrier)

```

4. Use the `one_of()` helper function to select the following columns -  year, month, day, dep_time, sched_dep_time, dep_delay, arr_time, from the flights data.

```{r}
CoList1 <- c("year", "month", "day", "dep_time", "sched_dep_time", "dep_delay", "sched_dep_time", "dep_delay", "arr_time")
flights %>% 
  select(one_of(CoList1))
#did I need to make a vector out of these to make this work?
```

5. Use the `slice()` function to slice the first 10 rows in the flights data.

```{r}
flights %>% 
  slice(1:10)
```

6. Perform the same operation as the previous task but this time use `filter()`. (Hint: Search for how to get the row numbers of a tibble)

```{r}
flights %>% 
filter(row_number() <= 10)
#https://stackoverflow.com/questions/25994307/filtering-data-frame-based-on-row-number
```

7. Use the `top_n()` command to identify the top 5 flights out of NYC that experienced the greatest arrival delays in 2013.
NOTE - all flights in nycflights are in nyc

```{r}
#arr_delay, origin == NYC

flights %>% 
  top_n(5, arr_delay) %>% 
  arrange(desc(arr_delay))

#test that prior code worked
flights %>% 
  arrange(desc(arr_delay))

```


8. Use `top_n()` along with `arrange()` and `slice()` to pull out the 50th most delayed departure from NYC in 2013 (Not all 50, just the 50th). 
```{r}
flights %>% 
  arrange(desc(dep_delay)) %>% 
  slice(50)
```

Test results

```{r results = 'hide'}
#test
flights %>% 
  arrange(desc(dep_delay))  
```


9. Do the same thing as the previous task using 
    + `row_number()`
    + and `dense_rank()`
```{r}
flights %>% 
  arrange(desc(dep_delay)) %>% 
  filter(row_number() == 50)
#this appears to be correct
```

```{r}
flights %>% 
  arrange(desc(dep_delay)) %>% 
  filter(dense_rank(desc(dep_delay)) == 50)
#test

```


10. The result from `row_number()` and `top_n()` are the same, while that from `dense_rank()` is different. Why?

    + Answer: because dense rank, according to my notes, eliminates ties, as in it takes the 
    first row it finds and then skips other rows with the same value while ignoring them for
    the rank count. I think top_n and row number include ties while getting to rank 50.

11. Find the top 10 airline carrier that had the highest average departure delays in 2013 using `group_by()`, `summarise()` and other functions you have learnt previously.
```{r}
flights %>% 
  group_by(carrier, year) %>% 
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(desc(delay)) %>% 
  top_n(10, desc(delay))
  
```

12. Use `group_by()` with `mutate()` to create a new variable called `comparativeDelay` which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the `time_hour` variable in the flights data). Store the result in a variable called `comparativeDelays`.
```{r}
#the cute picture you expect comes from #Summarise
Comparative_delays <- flights %>% 
  group_by(origin, time_hour) %>% 
  mutate(meanDelay = mean(dep_delay)) %>% 
  mutate(comparativeDelay = dep_delay - meanDelay) %>% 
  ungroup(.self) %>% 
  select(comparativeDelay)

Comparative_delays
#ive assumed here that what we're storing is just the comparative_delay column but Q13 makes me wonder
  
```

13. Now group the `comparativeDelays` tibble by carriers to print the top 10 airlines with the greatest average comparative delays.
```{r}
Comparative_delays2 <- flights %>% 
  group_by(origin, time_hour) %>% 
  mutate(meanDelay = mean(dep_delay)) %>% 
  mutate(comparativeDelay = dep_delay - meanDelay) %>% 
  ungroup(.self) %>% 
  group_by(carrier) %>% 
  #how can we verify success of group_by, ungroup(.self), etc?
  summarise(mean(comparativeDelay, na.rm = T)) %>% 
  top_n(10)


Comparative_delays2
  
```

14. Use `group_by()` with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows.
```{r}
flights %>% 
  group_by(origin) %>% 
  top_n(5, arr_delay)
```

15. The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why?

    + Answer: I think method from Question 11 is fine. Question 13 tries to normalize by airport and then summarize the average of each airport I think that sort of distorts the whole picture. Question 14 is wholly inppaopriate.
    
##Task set 3: R markdown

1. Modify the text below as per the prompt
     Underline me  
    **Make me bold**  
    > Make me a block quote  
    *Make me italic*  
    ~~Strike through me~~  
    
2. Create a level three header called "My level 3 header" under this line  

###My level 3 header

3. Modify the content below so that the bullet points (unordered list) render correctly

* Messy bullet points
+ one
+ two
+ three

4. Use chunk options to hide the output and the code from the chunk below in the previewed notebook.

```{r, results='hide', echo=FALSE}
print("Hide me and the code that created me")
```


