---
title: "Class 21 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

## Recap
1. Load the gapminder and tidyverse packages
```{r}
library(tidyverse); library(gapminder)


gapminder %>% unique(gdpPercap)
```

2. Use `max()`, `min()` and `quantile()` commands on the `gdpPercap` variable. Use the `$` sign to pull the column. Visually confirm that the max and min values are also there in the quantile values
```{r}
max(gapminder$gdpPercap)
min(gapminder$gdpPercap)
quantile(gapminder$gdpPercap)
```


```{r}
quantile(gapminder$gdpPercap) %>% str()
```

```{r}
rand_vec <- 1:4
names(rand_vec) <- c("a", "b", "c", "d")

str(rand_vec)
```


3. Repeat the same steps as above but this time use `pull()` from dplyr instead of `$`. Pipe users get extra points.

```{r}
max(pull(.data = gapminder, gdpPercap))
```

```{r}
pull(.data = gapminder, gdpPercap) %>% max()
```

```{r}
gapminder %>% pull(gdpPercap) %>% max()
```


4. Repeat the same steps as 2 but this time with `[[]]`.
```{r}
min(gapminder[["gdpPercap"]])
max(gapminder[["gdpPercap"]])
quantile(gapminder[["gdpPercap"]])
```

Will not work because filter needs a df.
```{r}
filter(gapminder$country == "India")
```


5. The default specification of `quantile()` gives the 0th (min), 25th, 50th (median), 75th and 100th (max) percentile values in a numeric vector. You can change these using the `probs` argument. Set probs to c(0.4, 0.6) to find the 40th and 60th percentile values of gdpPercap.
```{r}
quantile(gapminder$gdpPercap, probs = c(0.4, 0.6))
```


6. Show the most recent year of data for Afghanistan. You can find the most recent year by setting year to be equal to the max year value.

```{r}
gapminder %>% 
  filter(gapminder$country == "Afghanistan") %>% 
  filter(year == max(gapminder$year))
```
The lengths don't match because you are referring to a data which outside of the pipe that has not undergone the same transformations as the piped data.
```{r}
gapminder %>% 
  filter(year == max(gapminder$year)) %>% 
  filter(., gapminder$country == "Afghanistan") 
```

```{r}
bob_filter <- gapminder$country == "Afghanistan"
```


```{r}
gapminder %>% 
  filter(country == "Afghanistan" & year == max(year))
```

7. Follow the same logic to show the oldest row of data for Afghanistan.

```{r}
gapminder %>% 
  filter(country == "Afghanistan" & year == min(year))
```



8. Use `quantile()` and filter rows that are below the 2nd percentile value of gdpPercap in the most recent year of the data. If you are getting two rows in your output, think about the data that quantile sees when it calculates the 2nd percentile. The correct answer should have Burundi in it as well.
```{r}
quantile(gapminder$gdpPercap, probs = 0.02)
```

this will not work because quantile is being calculated on the entire dataset instead of the most recent year.
```{r}
gapminder %>% 
  filter(year == max(year) & gdpPercap < quantile(gdpPercap, probs = 0.02))
```


```{r}
gapminder %>% 
  filter(year == max(year)) %>% 
  filter(gdpPercap < quantile(gdpPercap, probs = 0.02))
```

```{r}
gapminder %>% 
  filter(year == max(year)) %>% 
  pull(gdpPercap) %>% 
  quantile(probs = 0.02)
```


9. Create a new column in the gapminder using `mutate()` that is the square root of gdp per cap. Call it anything you want. You can find the square root of a number as follows `100 ^ 0.5`.
```{r}
gapminder %>% 
  mutate(gdp_sr = gdpPercap ^ .5)
```

10. Filter the earliest year in the data and create a new column using `mutate()` that checks if `lifeExp` is below the median and `gdpPercap` is below the median. If yes, mark as "High mortality and poverty" else mark as "Other". Use an `if_else()` statement to do this. Keep only this new column along with country, year and lifeExp.
```{r}
oldest <- gapminder %>% 
  filter(year == min(year))

median_gdp <- quantile(oldest$gdpPercap, probs = 0.5)

median_lifeExp <- quantile(oldest$lifeExp, probs = 0.5)

oldest %>% 
  mutate(below_median = if_else(lifeExp < median_lifeExp & gdpPercap < median_gdp, "High Mortality", "Other")) %>% 
  select(-2, -5)
```


```{r}
gapminder %>% 
  filter(year == min(year)) %>% 
  mutate(below_median = if_else(lifeExp < median(lifeExp) & gdpPercap < median(gdpPercap), "High Mortality", "Other")) %>% 
  select(-2, -5)
```

```{r}
T & F
```


11. Filter the earliest year in the data and create a new column `gdpPercapLevel` that marks everything below the 25th percentile of gdpPercap as "Low GDP per cap", those above 75th percentile of gdpPercap as "High GDP per cap" and everything else as "Medium GDP per cap". Use either a nested `if_else()` or better a `case_when()`.

This uses a nested if_else.
```{r}
gapminder %>% 
  filter(year == min(year)) %>% 
  mutate(gdpPercapLevel = if_else(gdpPercap < quantile(gdpPercap, probs = 0.25),
                                  "Low GDP per cap", 
                                  if_else(gdpPercap > quantile(gdpPercap, probs = 0.75),
                                          "High GDP percap", 
                                          "Medium GDP per cap")))
  
```
Lets try this with a case_when() which is a bit easier to read.
```{r}
gapminder %>% 
  filter(year == min(year)) %>% 
  mutate(gdpPercapLevel = case_when(
    gdpPercap < quantile(gdpPercap, probs = 0.25) ~ "Low GDP per cap",
    gdpPercap > quantile(gdpPercap, probs = 0.75) ~ "High GDP percap",
    TRUE ~ "Medium GDP per cap"
  ))
```

