---
title: "Class 11 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
**LEARNING OBJECTIVE**: Subset data using base R syntax

**TOPICS TO COVER**

1. Subset atomic vectors in base R
    + Using `$`
    + Using `[]`
2. Subset data frames in R
    + Using `[]`

```{r}
intVect <- 1:100

intVect
```



##Subsetting atomic vectors
Lets say we want to extract the departure delay column from flights. 
```{r}
library(tidyverse); library(nycflights13)

depDelays <- flights$dep_delay
depDelays
```

Now I would like to see departure delays that are greater 540 mins
```{r}
depDelays <- depDelays[depDelays > 540]
```

What we need to do is remove these NA values as well in the our condition.
```{r}
depDelays <- depDelays[depDelays > 540 & !is.na(depDelays)]
```

I could combine all of these steps in one go.

```{r}
flights$dep_delay[flights$dep_delay > 540 & !is.na(flights$dep_delay)]

depDelays
```
Lets try this in tidyverse
```{r}
flights %>% 
    filter(dep_delay > 540) %>% 
    pull(dep_delay)
```

##Base subset of dataframes

Selecting columns
```{r}
flights[c(3, 4)]
```

```{r}
flights[3:4]
```

```{r}
flights[c("day", "dep_time")]
```
This will not work because we haven't added qoutes.
```{r}
flights[day, dep_time]
```

Lets now filter from a data frame.
```{r}
flights[flights$dep_delay > 540 & !is.na(flights$dep_delay), ]
```

This will not work
```{r}
flights[dep_delay > 540 & !is.na(dep_delay), ]
```

The last thing is to combine select and filter
```{r}
flights[flights$dep_delay > 540 & !is.na(flights$dep_delay), c(1:3, 5)]
```

The tidyverse version
```{r}
flights %>% 
    filter(dep_delay > 540) %>% 
    select(1:3, 5)
```













