---
title: "Class 2 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
**INSTRUCTIONS**:

1. Give your answers below each numbered task. Enter a new line.
2. The tasks are clustered under level 2 headers (`##`). Do one set at a time.
3. Preview your notebook at the end of each in class exercise task set. Does the notebook look formatted correctly? If not fix it. In most cases the issue is

**KEYBOARD SHORTCUTS**:

1) Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2) Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3) Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4) Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5) Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

#Class 3 Tasks
##Filter rows (Task time 15 mins)
1. Select the first three columns (year, month, day), all other columns whose names `contains()` the term "dep" and the origin column. Filter this data to show all the flights that took off in the morning (before 12:00) from JFK in December. Make sure to use pipes between the select and filter command. Refer to the shortcut for inserting a pipe (see above).
```{r}
flights %>% 
    select(1:3, contains("dep"), origin) %>%
    filter(dep_time < 1200 & origin == "JFK" & month == 12)
```
2. Use `top_n()` to print the most delayed departures from NYC in 2013. Read the documentation for `top_n()` on tidyverse if you are confused.
```{r}
flights %>% 
    top_n(., n = 5, wt = dep_delay)
```

3. Run the code below, read the error and fix the code so that it works
```{r}
flights %>% 
    filter(month == 6 & day > 15)
```
5. Filter the flights that were between the 10th and 40th most delayed in terms of arrival (`arr_delay`) using the `dense_rank()` helper function.
```{r}
flights %>% 
    filter(dense_rank(desc(arr_delay)) %in% 10:40)
```


##Mutate a tibble (Task time 10 mins)
1. Create a variable that indicates whether a flight took off in the AM or the PM.
2. Use `transmute()` instead of `mutate()` to do the same. What is the difference between the two?

##Grouped operations (Task time 30 mins)
1. Find the top 10 airline carrier that had the highest average departure delays in 2013 using `group_by()`, `summarise()` and other functions you have learnt previously.
2. Use `group_by()` with `mutate()` to create a new variable called `comparativeDelay` which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the `time_hour` variable in the flights data). Store the result in a variable called `comparativeDelays`.
3. Now group the `comparativeDelays` tibble by carriers to print the top 10 airlines with the greatest average comparative delays.
4. Use `group_by()` with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows.
5. The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why?

##ggplot
1. Make a scatterplot of your choice using any two numeric variables from the flights dataset


