---
title: "Class 3 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---

##Learning Objectives
1. ~~Host your notebook online using github~~
2. ~~Filter rows from tabular data~~
3. ~~Mutate tabular data~~
4. ~~Master grouping~~
5. Introduce ggplot2 for data visualizations

```{r, message=FALSE, warning=FALSE}
library(nycflights13); library(tidyverse)
```


##Filter rows (recap)
0. Get the unique list of carriers from flights
```{r}
unique(flights$carrier)
```
Getting unique/distinct values using tidyverse functions
```{r}
flights %>% 
    distinct(carrier) %>% 
    pull()
```


1. Filter out the flights that belong to the following carriers: UA, AA and DL
```{r}
flights %>% 
    filter(carrier == "UA" | carrier == "AA" | carrier == "DL" | carrier == "B6")
```
2. Use `%in%` to filter out the flights that belong to the following carriers: UA, AA and DL
```{r}
carriersToFilter <- c("UA", "AA", "DL", "B6")

flights %>% 
    filter(carrier %in% carriersToFilter) ## quasiqoutation / tidy evaluation
```

```{r}
flights
```


Sample for %in%.
```{r}
matchValue <- c(1, 2, 4, 5)
testVect <- 1:15

matchValue %in% testVect
```


3. Filter out the flight that suffered the highest departure delay in 2013.
```{r}
flights %>% 
    filter(dep_delay == max(dep_delay, na.rm = T))
```

Show missing dep delay values
```{r}
flights %>% 
    filter(is.na(dep_delay))
```


The reason for using na.rm = T
```{r}
testVector <- c(1, 2, 3, NA)

sum(testVector, na.rm = T)
```
4. Filter out the 50th most delayed flight that flew out from JFK in 2013 (using top_n() or arrange() + slice() and dense_rank())
Lets use top_n() first.
```{r}
flights %>% 
    top_n(., n = 50, wt = dep_delay) %>% 
    arrange(desc(dep_delay)) %>% 
    slice(50)
```

```{r}
flights %>% 
    filter(dense_rank(desc(dep_delay)) == 50)
```
5. Filter departure delays that are between one to two hours (show with and without helper)
```{r}
flights %>% 
    filter(dep_delay >= 60 & dep_delay <= 120)
```

```{r}
flights %>% 
    filter(between(x = dep_delay, left = 60, right = 120))
```


##Mutate
A simple trivial mutate.
```{r}
flights %>% 
    select(1:4) %>% 
    mutate(newDepTime = dep_time + 1200)
```


1. Use the lubridate package to create a new date variable and arrange it in descending order.

Introducing the paste function. It can paste together individual arguments like below,
```{r}
paste("R4D", "class", "is awesome", sep = "-")
```

or paste together elements in a vector. Notice that I have used `sep` as an argument in the first instance and `collapse` in the second.
```{r}
messageToCollapse <- c("R4D", "class", "is awesome")
paste(messageToCollapse, collapse = "-")

```

Now lets create new date variable using the `ymd()` function in the lubridate package.
```{r}
library(lubridate)

flights %>% 
    select(1:3) %>% 
    mutate(nycDates = ymd(paste(year, month, day, sep = "-")))
```


2. Use `if_else()` syntax to create a variable that indicates whether a flight was delayed.

```{r}
flights %>% 
    select(1:3, dep_delay) %>% 
    mutate(flightIsDelayed = if_else(dep_delay > 0, "Flight is delayed", "Flight is early or on time"))
```


##Grouping

1. Group and summarise
Example for grouping

```{r}
flights %>% 
    group_by(origin) %>% 
    summarise(avgDelay = mean(dep_delay, na.rm = T))
```

```{r}
flights %>% 
    group_by(carrier) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```

If you wanted to wanted see the groups along with other characteristics of a tibble use `glimpse()`
```{r}
flights %>%
    group_by(origin, carrier) %>% 
    glimpse()
```


```{r}
flights %>% 
    group_by(carrier)
```

Lets look at summarise without groups

```{r}
flights %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```

Lets summarise with two groups
```{r}
flights %>% 
    group_by(carrier, origin) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```




2. Counteracting data outliers in summary stats
Always look at the number of observations that a summary stat is based on.
```{r}
flights %>% 
    group_by(carrier, origin) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n()) %>% 
    arrange(nObs)
```

Add a few more summary functions to summarise()
```{r}
flights %>% 
    group_by(carrier, origin) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n(),
              maxDelay = max(dep_delay, na.rm = T),
              minDelay = min(dep_delay, na.rm = T),
              sd = sd(dep_delay, na.rm = T))
```


2. Group and mutate

```{r}
flights %>% 
    group_by(origin) %>% 
    mutate(meanDelay = mean(dep_delay, na.rm = T),
        comparativeDepDelay = dep_delay - meanDelay) %>% 
    select(1:3, origin,dep_delay, meanDelay, comparativeDepDelay)
```

3. Group and filter
```{r}
flights %>% 
    group_by(origin) %>% 
    filter(dep_delay == max(dep_delay, na.rm = T)) %>% 
    select(1:3, origin, dep_delay)
```

```{r}
flights %>% 
    filter(dep_delay == max(dep_delay, na.rm = T)) %>% 
    select(1:3, origin, dep_delay)
```


##A ggplot taster
1. Draw an exploratory chart
2. Explain the underlying concepts: data, aethetics, geoms and scales


```{r}
source("../helpers.R")

pickRandom()
```



