---
title: "Class 10 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
**LEARNING OBJECTIVE**: Explore data effectively using a checklist

**TOPICS TO COVER**

1. Next assignment: Explore the causes of departure and arrival delays

2. ~~Formulate your initial question~~
3. ~~Examine your data sources~~
    + ~~`str()`, `glimpse()`, `summary()`, `skim()`~~
    + ~~Check the top and bottom of your data~~
3. ~~Examine the variables of interest for your question~~
4. ~~Explore a solution for your question~~
    + Use visual comparisons
    + Illustrate causal frameworks
    + Use multivariate comparisons
5. Try and invalidate your solution
6. Reformulate your question (Rinse and repeat)
7. Integrate different modes of presenting evidence

##Formulate your initial question

We would like to explore the relationship between departure delay and weather?

There is a relationship between departure delays and weather

There is a relationship between departure delay and inclement weather

~~Which weather variable has the strongest impact on departure delay?~~

The last question in being specific is also disregarding other potential causes of departure delay.

##Examine your data sources
We have two data sources that we will be using - `flights` and `weather`. So first lets take a look at them.

Here is the flights data printed.
```{r}
library(tidyverse); library(nycflights13)

flights
```

You can also look at the structure of the flights data using the `str()` command. This gives us a bit more detail about the data.
```{r}
str(flights)
```

A similar command is `glimpse()` from `dplyr`.
```{r}
glimpse(flights)
```

We can also see basic summary statistics about the variables in our data using the `summary()` command.
```{r}
summary(flights)
```
Finally, the `skim()` function brings most of these functions together with a single command that neatly displays the characteristics of our data with histograms. Going forwards this would be the only command you would need to use when beginning to explore your data.
```{r}
library(skimr)

skim(flights)
```

Now lets look at the weather data.
```{r}
skim(weather)
```

Lets identify the primary keys of weather data. First guess `time_hour` should uniquely identify every obs.
```{r}
weather %>% 
    group_by(time_hour) %>%
    summarise(nObs = n())
```
This seems to indicate that the weather is being recorded for every origin.
```{r}
weather %>% 
    group_by(origin, time_hour) %>% 
    filter(n() > 1)
```

What are the keys for the flights data?

h/t to Mario and team the keys are as below.
```{r}
flights %>% 
    group_by(time_hour, flight, carrier) %>% 
    filter(n() > 1)
```

##Examine the variables of interest for your question
The first variable that we would like to explore is the dependent variable, which in this case is departure delay.
```{r}
summary(flights$dep_delay)
```

Now lets look at how it is distributed
```{r}
ggplot(data = flights, aes(dep_delay)) +
    geom_histogram()
```
This looks kinda meh, lets change the bins.

```{r}
ggplot(data = flights, aes(dep_delay)) +
    geom_histogram(bins = 100)
```
Another way to display a continuous variable is using a density plot
```{r}
ggplot(data = flights, aes(dep_delay)) +
    geom_density()
```

Lets look at a table as well. Here I cut (using `cut_width()`) the `dep_delay` variable and display the counts within. This is a table form of the histogram and is another way to get a sense for the spread of the data.
```{r}
flights %>% 
    filter(dep_delay > 0) %>% 
    mutate(binnedDelay = cut_width(dep_delay, width = 40)) %>% 
    group_by(binnedDelay) %>% 
    summarise(nObs = n())
```

Now lets look at the `wind_gust` variable and explore this further.
```{r}
ggplot(data = weather, aes(wind_gust)) +
    geom_histogram(bins = 60)
```

I am interested in seeing the unique values that wind gust takes. h/t mehrnoosh
```{r}
unique(weather$wind_gust)
```

##Explore a relationship
Let's create a cut of the flights data that we are interested in.

```{r}
delayedFlights <- flights %>% 
    filter(dep_delay > 0)
```


We would like to explore the relationship between departure delay and wind gust. Our first task is to join in the weather data.

```{r}
weatheredFlights <- delayedFlights %>% 
    select(origin, time_hour, dep_delay) %>% 
    left_join(x = ., 
              y = weather, 
              by = c("origin", "time_hour"))
```

Segue to a duplicate left join.

```{r}
left_join(x = weather, y = flights, by = c("origin", "time_hour"))
```

```{r}
weather
```



Next lets do a few simple plots.

```{r}
weatheredFlights
```

```{r}
weatheredFlights %>% 
    filter(dep_delay > 0) %>% 
    group_by(wind_gust) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))%>% 
    ggplot(aes(x = wind_gust, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm")
```

This chart looks interesting/funky. Lets explore the number of observations within each of these groups.

```{r}
weatheredFlights %>% 
    group_by(wind_gust) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n())
```

Lets consolidate some of the smaller groups and make a plot.
```{r}
plotData <- weatheredFlights %>% 
    filter(dep_delay > 0) %>% 
    mutate(wind_gust = if_else(wind_gust > 40, 40, wind_gust)) %>% 
    group_by(wind_gust) %>% 
    # ##setting a threshold of 20 obs within each group
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n())

ggplot(plotData, aes(x = wind_gust, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm")
```


Lets group by origin and the wind_gust variable. However, since we are looking at smaller groups lets also set a Nobs threshold
```{r}
plotData <- weatheredFlights %>% 
    mutate(wind_gust = if_else(wind_gust > 40, 40, wind_gust)) %>% 
    group_by(wind_gust, origin) %>% 
    # ##setting a threshold of 20 obs within each group
    summarise(meanDelay = mean(dep_delay, na.rm = T),
              nObs = n())
    



```

```{r, width=12}
ggplot(plotData, aes(x = wind_gust, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ origin, nrow = 1)
```

Hari's way

```{r, width = 12}
weatheredFlights %>% 
    group_by(origin, wind_gust) %>% 
    filter(n() > 20) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = wind_gust, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~origin)
```



```{r}
flights
```












