---
title: "Class 14 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

##Task Set 1: Load and compare
1. Load the messy gapminder data using `read_csv()`. If you saved the data inside a data folder that is in the same folder as this notebook the path would be: "data/gapminder_wide.csv".

2. Use `head()` to print out the first 10 rows (you will need to set the n argument within the function to 10 to change it from the default of 6)

3. Compare the output above with the `gapminder` data from the gapminder package.
```{r}
gapminder::gapminder
```

4. What are some issues with the wide version of the data (from the perspective of tidyness)


##Task Set 2: Plan
1. Discuss and list the issues with this data within your group (from the perspective of tidyness)

2. Write down the steps you will take to clean this data up (in terms of `tidyr` functions). So for instance: "I will gather xyz columns, then separate a_b_c...."

##Task Set 3: Clean
1. Load and clean the untidy gapminder data.
Here is my implementation for this.
```{r, message=FALSE}
library(tidyverse); library(gapminder)
gapminderWide <- read_csv("assets/gapminder_wide.csv")

gapminderWide
```

It looks like all the columns are encoding a particularly statistic and the year. Lets gather those into a combo column.
```{r}
gapminderWide <- gapminderWide %>% 
    gather(key = "combo", value = "value", -continent, -country)

gapminderWide
```

Next step we need to split the combo into year and variable.
```{r}
gapminderWide <- gapminderWide %>% 
    separate(col = combo, into = c("variable", "year"), sep = "_")

gapminderWide
```

Now the last task is to spread the variables into a column.
```{r}
gapminderWide <- gapminderWide %>% 
    spread(key = variable, value = value)

gapminderWide
```

Lets compare the output with the actual gapminder data.
```{r}
gapminder
```

The order of columns is not the same across the two tibbles. Lets fix that by reordering the gap_wide data by the names of the actual gapminder data.

```{r}
gapminderWide <- gapminderWide %>% 
    select(names(gapminder))

gapminderWide
```

Bonus task for everyone make the column types also match.
