---
title: "Class 8 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
**LEARNING OBJECTIVE**: Revise content that we have covered so far.

**Topics to cover**

1. Common errors
    + Syntax errors
    + Conceptual errors
2. Groups


##Common Syntax errors

Incomplete expression
```{r}
library(tidyverse); library(nycflights13)

flights %>% 
    group_by(origin) %>% 
    summarise(
        meanDelay = mean(dep_delay, na.rm = T),
        maxDelay = max(dep_delay, na.rm = T)
```

```{r}
flights %>% 
    group_by(origin) %>% 
    summarise(meanDelay = mean(dep_delay), na.rm = T)
```

The way to correct the code is below
```{r}
flights %>% 
    group_by(origin) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```
This is an incorrect assignment pattern.
```{r}
flights %>% 
    filter(dep_delay > 0) <- delayedFlights
```

The correct assignment pattern is as follows
```{r}
delayedFlights <- flights %>% 
    filter(dep_delay > 0)
```

This is not advised
```{r}
flights %>% 
    filter(dep_delay > 0) -> delayedFlights2
```

This will not work when I rerun this file because I have created the object after the line of code that is using it.
```{r}
sum(x)
x <- 1:100
```

##Conceptual errors

```{r}
sum(x, narm = T)
```

Resetting .data argument within a set of pipes.
```{r}
flights %>% 
    filter(dep_time < 1200) %>% 
    filter(.data = flights, dep_delay > 0) 

%>% 
    group_by(origin) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```

Another 

```{r}
flights %>% 
    filter("carrier" == "AA")
```

```{r}
flights %>% 
    filter("carrier" > "AA")
```

WAT!?



##Lets revise grouping
```{r}
flights %>% 
    filter(carrier %in% c("AA", "DL", "9E", "EV")) %>% 
    group_by(month) %>% 
    sample_n(size = 2) %>% 
    select(1:3, dep_delay, carrier) %>% 
    group_by(month, carrier) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T))
```







