---
title: "Class 12 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---

**Learning Objective 1:** Understand the concept of vectorization

**Learning Objective 2:** Read and write data from different formats into R

**Topics**

1. Vectorization
2. Difference between summarising functions and other (say sum vs log)
3. == vs. %in%
4. Working directories
5. Navigation within R
6. Loading data into R
7. Writing data into R
8. Bonus: Working with google drive

```{r}
x <- 1:3
y <- 1:6

```

```{r}
x + y
```

```{r}
z <- 1:7
z
```

```{r}
x + z
```


```{r}
x == y
```


```{r}
c(1, 2, 3) == c("1", "2", "3")
```

If you want to check if two objects in R are identical...
```{r}
identical(x, y)
```

```{r}
x + y
sum(x, y)
```


```{r}
sum(x + 2)
```


```{r}
sum(x) + 2
```


I would like to illustrate the difference between using `==` vs `%in%` operator.

```{r}
x == y
```

```{r}
x
```

```{r}
y
```

```{r}
x %in% y
```
I would like to pull out the carriers "UA", "AA", "B6"
Why is this the wrong way to do this?
```{r, message=FALSE}
library(tidyverse); library(nycflights13)

flights %>% 
    filter(carrier == c("UA", "AA", "B6"))
    
```


```{r}
flights %>% 
    filter(carrier %in% c("UA", "AA", "B6"))
```
What happens if we swap the order of operands.
```{r}
flights %>% 
    filter(c("UA", "AA", "B6") %in% carrier)
```

There is a place for the `==`.
```{r}
flights %>% 
    filter(dep_delay == 2)
```

```{r}
flights %>% 
    filter(dep_delay %in% 2)
```


##File traversal in R
You can load csv using base r.
```{r}
airlinesCSV <- read_csv(file = "data/airlines.csv")
```

```{r}
list.dirs()
```

I can also get the list of files in a directory
```{r}
list.files()
```

```{r}
list.files(pattern = "*.Rmd")
```


```{r}
library(readxl)

nycFlightsXL <- read_excel(path = "../outside data folder/nycFlightsData.xlsx", sheet = "weather")
```
If you have spss, stata, sas,
```{r}
library(haven)

airlinesDTA <- read_dta("data/airlines.dta")
```


```{r}
getwd()
```


##Linking with googlesheets
Current package to use is googlesheets. However, this will be deprecated soon (because of changes in the Google API) and be replaced by a combination of two packages googledrive and googlesheets4. However, if you learn googlesheets the syntax and concepts are similar to those in the packages that will eventually replace it

```{r}
library(googlesheets)

##register a spreadsheet
attendance <- gs_title("Attendance List")

##load a worksheet from a registered spreadsheet
attendanceData <- attendance %>% 
    gs_read(ws = "Attendance Sheet")

```

delete a sheet

```{r}
##delete a worksheet
attendance %>% 
    gs_ws_delete(ws = "temp")
```


```{r}
newAttendanceData <- attendanceData %>% 
    mutate(testCol = "hi")
```

```{r}
##add a new sheet
attendance %>% 
    gs_ws_new(ws_title = "temp", input = newAttendanceData, trim = TRUE, verbose = FALSE)
```


You can add a row

You can go in and edit specific cells.
