---
title: "Class 16 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

**Learning Objective**: Write a for loop and understand the basic recipe for it.

###Topics
1. Revise conditionals
2. For loops
3. Introduce the map family


##Zoom poll

Question 1 on the poll
```{r}
x <- 5
y <- 4

y %% 2 == 0 ## is y even?
```

```{r}
if(condition){
    Do this
} else {
    Do this
}
```


```{r}
if(y %% 2 == 0) {
    x <- x - 1
} else {
    x <- x + 1
}

x
```

When there is an error in the condition nothing gets run and code is broken.
```{r}
if(z %% 5 == 0) {
    x <- x - 1
} else {
    x <- x + 1
}
```

Question 3
```{r}
x <- c(6, 5)

if(x %% 5 == 0) {
    x <- x - 1
} else {
    x <- x + 1
}
```
if else block syntax is not vectorized. 
```{r}
x
```

the basic syntax for if_else
```{r}
if_else(condition, val1, val2)
```


```{r}
x <- c(6, 5)

x <- if_else(x %% 5 == 0, x - 1, x + 1)
```

```{r}
x
```

```{r}
x <- c(6, 5)
x <- if_else(x %% 5 == 0, "yes", "no")
```


What does `any()` do?

```{r}
x <- c(6, 5)

x == 5
```

any takes a vector and asks the question is there any TRUE values in you?

```{r}
any(x == 5)
```

```{r}
if(any(x == 5)) {
    x <- x + 1
} else {
    x <-  x - 1
}
```

```{r}
x
```

The values that true and false conditions take here would be either [7, 6] or for false [5, 4]
```{r}
x <- c(6, 5)

x <- if_else(any(x == 5), x + 1, x - 1)
```

```{r}
x <- if_else(any(x == 5), 1, x - 1)
```

```{r}
any(x == 5)
```


Lets play around with this
```{r}
x <- c(6, 5)

x <- if_else(x == 5, x + 1, x + 2)
```

```{r}
x
```

```{r}
x <- c(6, 5, 10, 5, 6)

x <- if_else(x == 5, "yes equal to 5", "no not equal to 5")

x
```

```{r}
x == 5
```

```{r}
any(x == 5)
```


```{r}
x <- c(6, 5)

x <- if_else(any(x == 5), x + 1, x - 1)
```

```{r}
x <- if_else(any(x == 5), "yes equal to 5", "no not equal to 5")

x
```

##For loops

```{r}
fruits <- c("apples", "oranges", "pineapple")

for (emma in fruits) {
    print(emma)
}
```

For every column in this dataset I would like to take its mean if it is numeric
```{r}
library(nycflights13)


flights
```

Lets get the names of the columns
```{r}
col_names <- names(flights)

col_names
```

```{r}
col_names[13]
```


Pull each column out and check if it is numeric.
```{r}
pull(flights, col_names[13]) %>% is.numeric()
```
if I wanted to repeat it for 1
```{r}
pull(flights, col_names[1]) %>% is.numeric()
```

```{r}
pull(flights, col_names[4]) %>% is.numeric()
```

```{r}
col_names
```

```{r}
1:length(col_names)
```

```{r}
1:0
```


```{r}
seq_along(col_names)
```

```{r}
for(i in seq_along(col_names)) {
    print(i)
}
```
```{r}
col_names
```


```{r}
for(i in seq_along(col_names)) {
    print(col_names[i])
}
```

```{r}
col_names
```



```{r}
for(i in col_names) {
    print(i)
}
```


Lets say I want to store the output insteadf printing it
```{r}
output <- vector(mode = "character", length = length(col_names))

for(i in seq_along(col_names)) {
    output[i] <- paste(col_names[i], "is the", i, "column")
}

output
```
```{r}
flights
```


```{r}
output <- vector(mode = "numeric", length = length(col_names))

for(i in seq_along(col_names)) {

    ##pull the ith column
    col_data <- pull(flights, )

    
    ##check if it is numeric
    if(is.numeric(col_data)) {
        ##store output
        output[i] <- mean(col_data, na.rm = T) %>% round(digits = 2)
    } else {
        output[i] <- 1
    }
    
}

output
```


After class

THere are two types of subsets in R. Simplifying vs preserving.
```{r}
flights %>% 
    select(1)
```

```{r}
flights %>% 
    pull(1) %>% 
    str()
```

The base R version of select
```{r}
flights[1] %>% str()
```


```{r}
flights[[1]] %>% str()
```

More commonly
```{r}
flights$year %>% str()
```

```{r}
is.numeric(c(NA, 1, 2, 3))
```




