---
title: "Class 6 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---

##Task Set 1: Practice Practice Practice
**Time**: 30 minutes

1. Install the 'gapminder' package. If you are using the `install.packages()` function to do this, make sure to delete the function once the package is installed.
2. Why should you delete the `install.packages()` function from your notebook once you run it?
3. Use `?gapminder` to access the help file for the dataset called `gapminder` from the gapminder package. Copy and paste the description of the dataset below.
4. Use `head()` to print the first 10 rows of the gapminder dataset.
5. Use `tail()` to print the last 20 rows of the gapminder dataset.
6. Use `str()` to see the structure of the gapminder dataset.
7. Use `glimpse()` to see the structure of the gapminder dataset.
8. Use `names()` to print the names of the columns in the gapminder dataset
9. Draw a histogram of the `gdpPercap` variable in the gapminder dataset.

```{r}
names(gapminder)
```

```{r}
library(tidyverse); library(gapminder)

ggplot(data = gapminder, mapping = aes(gdpPercap)) +
    geom_histogram()
```


10. Draw a histogram of the `gdpPercap` but this time add a binwidth argument to the histogram. Use a value of your choice.
```{r}
ggplot(data = gapminder, mapping = aes(gdpPercap)) +
    geom_histogram(binwidth = 1000)

ggsave("allYears.png")
```


11. Filter the gapminder dataset to show only the latest year for each country. (Hint: you will need to group the data and filter for the most recent year using year == `max(year, na.rm = T)`, figure out the variable you should group by.)

This is a slightly more elaborate way to write this.
```{r}
gapminder %>% 
    group_by(country) %>% 
    mutate(minYear = min(year, na.rm = T)) %>% 
    filter(year == minYear)
```

The better way to write this.

```{r}
gapminder %>% 
    group_by(country) %>% 
    filter(year == max(year, na.rm = T))
```

12. Do the same steps as in 11 but this time pipe to a ggplot function that plots a histogram.
```{r}
gapminder %>% 
    group_by(country) %>% 
    filter(year == max(year, na.rm = T)) %>% 
    ggplot(aes(gdpPercap)) +
    geom_histogram(binwidth = 1000)

ggsave("mostRecentYear.png")
```

13. Observe the differences in the histogram from 12 and 10.
14. Find the average population every year and show that on a column chart
```{r}
gapminder %>% 
    group_by(year) %>% 
    summarise(meanPop = mean(pop, na.rm = T))
```

15. Find the average population every year and show that on a line chart
```{r}
gapminder %>% 
    group_by(year) %>% 
    summarise(meanPop = mean(pop, na.rm = T)) %>% 
    ggplot(aes(x = year, y = meanPop)) +
    geom_line()
```

