---
title: "Class 4 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---

**CHEATSHEETS**

1. [Data visualization](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)
2. [Data Manipulation](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)

**KEYBOARD SHORTCUTS**:

1. Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2. Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3. Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4. Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5. Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

##Task Set 1: Recap + Factors
1. Filter all departure delays that were between 10 and 20 minutes from LGA in 2013

```{r}
library(nycflights13)
library(tidyverse)

flights %>%
  filter(origin == 'LGA',dep_delay > 10 & dep_delay < 20) %>%
  select(1:3,dep_delay,origin)

```



2. Use `group_by()` and `filter()` to find the flights that had the greatest delay for each month in 2013.

```{r}
flights %>%
  group_by(month) %>%
  filter(dep_delay == max(dep_delay,na.rm = T)) %>%
  select(1:3,dep_delay) %>%
  arrange(month)
```


3. Write two lines on why it is important to ungroup a grouped dataset (once your are done with the grouping operations).

Because future operators will only work on your groups, rather than the whole data set

4. Look at the documentation for `weekdays()`. Use it on the `time_hour` column inside a `mutate()` to create a new column called `dayName` that records the day of the week that a flight took off. Save this new dataset as `flightsWeek`.


```{r}
flightsWeek <- flights %>%
  mutate(dayName = weekdays(time_hour,abbreviate = T))
  
  
```


5. Convert this into a factor variable called `fctDayNames` using `factor(flightsWeek$dayName)`. Display it using `table(fctDayNames)`

```{r}
fctDayNames <- factor(flightsWeek$dayName)
table(fctDayNames)
```


6. Use `levels()` on `fctDayNames` to order the levels of the factors to the following order: c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"). Show the output using `table(fctDayNames)`

```{r}
fctDayNames <- factor(flightsWeek$dayName,levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
table(fctDayNames)

```


7. Now combine the previous two steps in a `factor()` function by setting the value of the levels argument within it, also set the labels argument to c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"). Save this as `cleanFctDayNames` See `?factor`. Show output using `table(cleanFctDayNames)`. Observe the differences in output between 5, 6 and 7.

```{r}
cleanfctDayNames <- factor(flightsWeek$dayName,levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"),labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
table(cleanfctDayNames)

```


##Task Set 2: Create simple charts
1. Create a histogram of arrival delay.

```{r}
ggplot(data = flights, mapping = aes(arr_delay))+geom_histogram()
```


2. Change the statistic for the histogram from the default of "count" to "density" and plot again.

```{r}
ggplot(data = flights, mapping = aes(arr_delay))+geom_histogram(stat = 'density',na.rm = T)
```

3. Run the following code to create a stacked column chart. Look through the documentation and examples for `geom_col()` on the tidyverse portal to fix the code below and create a chart where the bars for each origin are not stacked.

```{r}
flights %>% 
    group_by(origin, month) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = factor(month), y = meanDelay, fill = origin)) +
    geom_col() +
  

```{r}
library(tidyverse); library(nycflights13)
flightsWeek <- flights %>% 
    mutate(dayName = weekdays(time_hour))

```
5. Convert this into a factor variable called `fctDayNames` using `factor(flightsWeek$dayName)`. Display it using `table(fctDayNames)`.
```{r}
fctDayNames <- factor(flightsWeek$dayName)

table(fctDayNames)
```

6. ~~Use `levels()` on `fctDayNames` to order the levels of the factors to the following order: c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"). Show the output using `table(fctDayNames)`~~ INCORRECT USE OF LEVELS FUNCTION.
7. Now combine the previous two steps in a `factor()` function by setting the value of the levels argument within it, also set the labels argument to c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"). Save this as `cleanFctDayNames` See `?factor`. Show output using `table(cleanFctDayNames)`. Observe the differences in output between 5, 6 and 7.

```{r}
cleanFctDayNames <- factor(flightsWeek$dayName, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"), labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))

table(cleanFctDayNames)
```


