---
title: "Class 4 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---

**CHEATSHEETS**

1. [Data visualization](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)
2. [Data Manipulation](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)

**KEYBOARD SHORTCUTS**:

1. Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2. Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3. Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4. Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5. Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

##Task Set 1: Recap + Factors
1. Filter all departure delays that were between 10 and 20 minutes from LGA in 2013
2. Use `group_by()` and `filter()` to find the flights that had the greatest delay for each month in 2013.
3. Write two lines on why it is important to ungroup a grouped dataset (once your are done with the grouping operations).
4. Look at the documentation for `weekdays()`. Use it on the `time_hour` column inside a `mutate()` to create a new column called `dayName` that records the day of the week that a flight took off. Save this new dataset as `flightsWeek`.
5. Convert this into a factor variable called `fctDayNames` using `factor(flightsWeek$dayName)`. Display it using `table(fctDayNames)`
6. Use `levels()` on `fctDayNames` to order the levels of the factors to the following order: c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"). Show the output using `table(fctDayNames)`
7. Now combine the previous two steps in a `factor()` function by setting the value of the levels argument within it, also set the labels argument to c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"). Save this as `cleanFctDayNames` See `?factor`. Show output using `table(cleanFctDayNames)`. Observe the differences in output between 5, 6 and 7.

##Task Set 2: Create simple charts
1. Create a histogram of arrival delay.
```{r}
ggplot(flights, aes(arr_delay)) +
    geom_histogram()
```

2. Change the statistic for the histogram from the default of "count" to "density" and plot again.
```{r}
ggplot(flights, aes(arr_delay)) +
    geom_histogram(stat = "density")
```

3. Run the following code to create a stacked column chart. Look through the documentation and examples for `geom_col()` on the tidyverse portal to fix the code below and create a chart where the bars for each origin are not stacked.
```{r}
flights %>% 
    group_by(origin, month) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = factor(month), y = meanDelay, fill = origin)) +
    geom_col(position = "dodge")
```
4. Read the documentation for `geom_point()` to create a scatter plot that compares arrival delay with departure delay. How would you describe the relationship between the two variables?
5. Use the chunk below to group the data by `time_hour` and calculate the mean departure delay and pipe to a ggplot to create a line chart showing `time_hour` on the x-axis and the mean departure delay on the y-axis. Can you identify the spikes in the average delay? Which parts of this would you zoom into to get a better sense for the average delay? 
```{r, fig.width=12}

```

##Task Set 2: Facets and multiple geoms
1. Create a facetted version of any of the two histograms you made in the previous task set. You can pick a facet of your choice (remember that a facet should be a categorical variable)
2. The code below is the same as task 3 in the previous task set. Modify it to create a facetted version of the chart (instead of the stacked bars).
```{r}
flights %>% 
    group_by(origin, month) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(aes(x = factor(month), y = meanDelay, fill = origin)) +
    geom_col() 
```

3. Extend the scatter plot you made in the previous section by adding a `geom_smooth()` to it that uses a linear model (method = "lm").

##Task Set 3: Make a pretty plot
1. Take any plot that you made in any of the previous set of tasks and
    + Add axis labels and chart title
    + Any additional aesthetics to make it prettier
    + Choose a theme template (either from ggthemes or from ggplot)
    + Make any additional changes you would like to make it more attractive
2. Save this plot using `ggsave()` as a png and share it on the class4 channel.




