---
title: "Class 2 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---
##Packages Used
1. tidyverse. Specifically dplyr from within the tidyverse
2. nycflights13 a dataset containing all the flights that took off from NYC in 2013.
3. lubridate a tidyverse package to handle dates in R

##Learning Objectives
1. ~~Use slack to ask questions more effectively~~
2. ~~Understand material from class 1~~
3. ~~Install and load packages~~
4. ~~Use the tidyverse portal to access information~~
5. ~~Explain the use of pipes and the dot symbol~~
6. ~~Arrange/sort a table~~
7. ~~Select columns from tabular data~~




##Review material from class 1
1. Variable assignment
```{r}
x <- c("a", "b", "c")
```
2. Data types: Create three different data types
```{r}
charVec <- x
numericVec <- 5:20
numSeq <- seq(from = 0, to = 50, by = 5)
logicalVec <- c(T, F, T, T, T, F)
```
3. Data structures: Create a vector, a list and a data.frame
```{r}
sampleList <- list(charVec, numericVec, numSeq, logicalVec)
length(sampleList)
```

```{r}
sampleDF <- data.frame(x = 10:30, y = 30:50, z = 10:30)
```


##Installing and loading packages
```{r}
library(tidyverse); library(nycflights13)
```



##Arrange a table
1. Introduction to the nycflights13 package
```{r}
flights
```
2. Introduction to tibbles
```{r}
str(flights)
```
3. Arrange the flights data using the departure delay variable
```{r}
flights %>% 
    arrange(desc(dep_delay))
```

4. Arrange the flights data using two different variables
```{r}
flights %>% 
    arrange(desc(month), arr_delay)
```

##Select columns from a tibble
1. Using columns numbers
```{r}
flights %>% 
    select(1, 4, 6)
```
2. Remove using negative indices
```{r}
flights %>% 
    select(-(c(4, 5, 7)))
```

3. Select by name
```{r}
flights %>% 
    select(year, month, day, dep_delay)
```

4. Using name range
```{r}
newData <- flights %>% 
    select(year:arr_time)
```

##Filter
1. Introduce all the relational and logical operators in R
2. Filter by a numeric value
```{r}
flights %>% 
    filter(month == 12)
```
3. Filter by a character value
```{r}
flights %>% 
    filter(origin == "JFK")
```

5. Use logical operators to combine filter conditions
```{r}
flights %>% 
    select(1:3, origin, dep_delay) %>% 
    filter(month == 12 & day == 31 & origin == "JFK")
```




