---
title: "Class 7 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
css: /Users/hsubhash/Documents/R4D/Assets/custom.css
---

**LEARNING OBJECTIVES**: Make beautiful ggplot charts using default features and extensions.

**TOPICS**

0. Housekeeping
    + Thoughts on the class prep
    + Asking questions
2. Facets and groups
3. Make beautiful charts using
    + Aesthetics (alpha, colour etc.)
    + Scales
    + Theme
    + Labels and annotations
4. Save a plot
5. Discuss extensions and helpers
    + [ggthemes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
    + [gganimate](https://gganimate.com/)
    + [ggforce](https://www.data-imaginist.com/2019/the-ggforce-awakens-again/)
    + [ggrepel](https://github.com/slowkow/ggrepel)
    + [scales](https://scales.r-lib.org/)
    + [Many more](https://www.ggplot2-exts.org/gallery/)
    
    
2. Facets and groups
```{r}
flights %>% 
    group_by(origin, hour) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) 
```

```{r}
flights %>% 
    filter(hour %in% 1:4)
```

```{r}
mean(NA, na.rm = T)
```
```{r}
sum(NA, na.rm = T)
```


```{r}
library(nycflights13); library(tidyverse)

flights %>% 
    group_by(origin, hour) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T)
```

Lets truncate this axis so that it doesn't show hours before 5. One way to do this is to filter the data.
```{r}
flights %>% 
    group_by(origin, hour) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ungroup() %>% 
    filter(hour != 1) %>% 
    ggplot(mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T)
```

Another way to do it is to explicitly set the scales.
```{r}
flights %>% 
    group_by(origin, hour) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(5, 24))
```

Lets now add a theme from the ggthemes package. I have also set the max value so that is dynamic.
```{r}
library(ggthemes)
plotData <- flights %>% 
    group_by(origin, hour) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ungroup()

ggplot(data = plotData, mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(5, max(plotData$hour, na.rm = T))) +
    theme_fivethirtyeight()

```

Fivethirtyeight removes the axis titles. Lets add those back.

```{r}
ggplot(data = plotData, mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(5, max(plotData$hour, na.rm = T))) +
    theme_fivethirtyeight() +
    theme(
        axis.title = element_text()
    )
```
Lets add some titles, titles, titles
```{r}
ggplot(data = plotData, mapping = aes(x = hour, y = meanDelay, colour = origin)) +
    geom_line(na.rm = T) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(5, max(plotData$hour, na.rm = T))) +
    labs(x = "Hour of departure",
         y = "Mean departure delay",
         title = "Don't fly late") +
    theme_fivethirtyeight() +
    theme(
        axis.title = element_text()
    )

ggsave("testPlot.png")
```

```{r}

```

