---
title: "Practice makes perfect"
author: Your Name
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

#Instructions {.tabset .tabset-fade}
This notebook contains practice material. The aim is to get you to revise and practice some of the concepts that we have learnt so far.

##Subset
###Setup
While we have used the `dplyr` package to extract rows and columns from tibbles, it is useful to keep the base R syntax for this handy. I will use the `gapminder` dataset for this section. Lets load it along with the tidyverse. Lets use the `skim()` function from `skimr` to look at the gapminder data before we go further.

```{r, message=FALSE}
library(gapminder); library(tidyverse); library(magrittr)
skimr::skim(gapminder)
```

Now let's practice.

###Pulling columns
There are multiple ways in which you can pull columns in base R. The first is using the `$` notation. Lets use this to select the continents column from gapminder.

```{r}
continents <- gapminder$continent
```

The analogous command in `dplyr` uses the `pull()` as follows. 
```{r}
continents <- pull(gapminder, continent)
```

Notice that I am using `pull()` instead of `select()`. This is because the output of a pull command or a `$` applied on a data frame or a tibble is an atomic vector while that from a `select()` is a tibble. See the output of the `str()` command below. It shows a factor vector (which is still an atomic vector) and not a data frame/tibble.
```{r}
str(continents)
```


I can also pull the continents using the `df[["columnName"]]` syntax. Lets try that.
```{r}
continents <- gapminder[["continent"]]
```

**TASK 1**: You can also use `df[[colNum]]` syntax to pull out a column from a tibble/dataframe. Try that below to complete the code.
```{r}
continents <- "Use an integer representing column position to pull out a column"
```

**TASK 2**: Now I would like you to either of these methods along with the `unique()` command to print the unique names of continents in the gapminder data. Replace the prompt with your code to do that. Don't use the continents variable that I have created before to complete this task.
```{r}
unique("ENTER YOUR CODE TO PULL COLUMN HERE")
```

**TASK 3**: Use the `table()` command the same way that you tried the `unique()` to show the number of observations for each continent in gapminder.
```{r}

```

**TASK 4**: The continent column in the gapminder dataset is currently stored as a factor. Use `as.character() %>% unique()` to first convert the vector to character and then find its unique values. Notice the difference in the output from when you compare the output with the earlier `unique()` command.
```{r}

```

**TASK 5**: Do the same thing as above without the pipe and without assigning to a variable.
```{r}

```

###Selecting a column
You can select a column using a single instead of a double bracket. See below.
```{r}
continentDF <- gapminder["continent"]
```

As noted earlier, the `select()` command returns the column as a data frame. See below.
```{r}
str(continentDF)
```

**TASK 6**: Identify the `unique()` values of continent, however, this time use the single bracket notation. Notice the difference in the output this time.
```{r}

```

**TASK 7**: Achieve the same thing as above using the `dplyr` function `distinct()`
```{r}

```

Another useful trick to specify the index of the columns you don't want in a tibble. You can do this by adding a minus sign next to the index. Take a look at the gapminder data below. Let's say I want to create a version of the gapminder data that does not include life expectancy or gdp per capita. 

```{r}
head(gapminder)
```

I can do this using the negative index.
```{r}
gapminder[c(-4, -6)] %>% 
    head()
```

You can do the same thing within a `select()` function in `dplyr`. However, since the tidyverse allows you to unquote the names of columns within a tibble you can also apply a negative sign next to the unquoted name to remove them from a tibble. 

```{r}
gapminder %>% 
    select(-lifeExp, -gdpPercap) %>% 
    head()
```
The same thing is not possible with base R and will throw and error since the names of the columns need to be specified as a character vector.
```{r}
gapminder[-c("lifeExp", "gdpPercap")] 
```

**TASK 8**: Remove the population and gdp per capita columns from gapminder data using base R syntax. Use the head command to only display the first 6 rows of this data.

**TASK 9**: Repeat the same task with `dplyr` syntax

A common task in data science is to use a vector of column names to extract them from a data set. You can get the names of the columns in a data frame using the `names()` function.
```{r}
names(gapminder)
```

Let's use that to extract the first to recreate the population dataset that I created earlier.
```{r}
gapminder[names(gapminder)[c(-4, -6)]] %>% 
    head()
```
Thats a complex and difficult to read piece of code so lets break that down a little. Let's do it from the inside out starting with `c(-4, -6)`. That command creates the list of columns I would like to remove which are the 4th and 6th columns in gapminder.
```{r}
colsToRemove <- c(-4, -6)
colsToRemove
```

Next lets look at the next piece `names(gapminder)`. As we saw earlier, this function gives the names of the columns in gapminder. Let's store that in a character vector as well.
```{r}
gapminderCols <- names(gapminder)
gapminderCols
```

Now, lets combine the two pieces that we just constructed to identify the names in gapminder that I am interested in selecting. The code below takes the character vector containing the names of the columns in gapminder and removes the values at the 4th and 6th index (lifeExp and gdpPercap). The output is the list of columns we are interested in keeping.
```{r}
colsToKeep <- gapminderCols[colsToRemove]

colsToKeep
```
Now lets use the list of columns to keep to subset the gapminder data.
```{r}
gapminder[colsToKeep] %>% 
    head()
```

The same task in `dplyr` would be achieved as follows.
```{r}
gapminder %>% 
    select(one_of(colsToKeep))
```
**A small segue into scoping**: The select command when it encounter `colsToKeep` looks for an object by that name in the global environment and since in this case it exists in the global environment it uses that to subset the data. If it did not exist in the global environment it would only then look for it within the piped data frame itself. If for some reason it does not exist in the piped data frame, it will start looking in the different loaded packages to see if it exists within any of those. If it doesn't then it will throw and error.

**TASK 10**: Use the base R syntax along with the `names()` to remove the population and gdp per capita variables from the gapminder data.
```{r}

```

###Filter rows
You can extract rows from a data frame in several ways in base R. The first way is to use a number specifying the position of a row. The code below slices the first row out from the gapminder data using an index. Notice that the syntax is slightly different from what we used to pull and select in the earlier sections. Instead of a single value within the square bracket notation there is now space for two that are separated by a comma. The syntax can be illustrated as follows - `df[rows, columns]`. The part before the comma is used to subset rows while that after the comma is used for columns. In our case, we would like to slice the entire first row. So we have left the subset command for columns empty which selects all of them.
```{r}
gapminder[1, ]
```

The equivalent in `dplyr` is as follows.
```{r}
slice(.data = gapminder, 1)
```
or
```{r}
gapminder %>% 
    slice(1)
```

You can also specify a range of rows to slice. The code below gives first to the 10th row in the data.
```{r}
gapminder[1:10, ]
```

**Task 11**: Use `seq()` to create a vector of even numbers from 2 to 20. Use this vector to slice the even rows until 20 (including 20). Use the base syntax for this.
```{r}

```

**Task 12**: Use `seq()` to create a vector of even numbers from 2 to 20. Use this vector to slice the even rows until 20 (including 20). Use the `slice()` function for this.

**Note**: We learnt earlier that we can use either the `$` (dollar) or the `[[]]` (double square bracket) operators to perform a simplifying subset i.e. take a data frame and pull out a column from it as an atomic vector. This is not possible for a row i.e. we cannot take a single row from a dataframe using either of these notations and covert it to an atomic vector. There are other ways in which you could do this, however, in general this is not advisable since you would be coercing a lot of different columns that hold potentially data of different types into an atomic vector with a single data type.

In most cases, we will use some sort of condition to filter rows. For instance, lets say that we only want the years greater than 1990 from out gapminder data. We can do this as follows.

```{r}
gapminder[gapminder$year > 1990, ] %>% 
    head()
```

Let's break this down a little more. `gapminder$year > 1990` returns a vector of TRUE or FALSE values indicating whether the atomic vector `year` that we pulled from the gapminder data using the `$` notation is greater than 1990. Below I am showing the first 10 values from this TRUE or FALSE vector.

```{r}
yearsGreaterThan1990 <- gapminder$year > 1990
yearsGreaterThan1990[1:10]
```
This vector has the same length as the number of rows in the gapminder data (since it is using the year column from the gapminder data to generate itself).
```{r}
paste("The number of rows in the gapminder data is", nrow(gapminder))
paste("The length of the logical vector of years greater than 1990 is", length(yearsGreaterThan1990))
```
This vector of TRUE or FALSE values is then used to subset the rows using the `df[rows, columns]` syntax. See below.
```{r}
gapminder[yearsGreaterThan1990, ] %>% 
    head()
```
This is the same as the code earlier, except that we created an intermediate variable to store the logical values that are the output of the comparison `gapminder$year > 1990`.

The same thing in `dplyr` can be done as follows.
```{r}
filter(.data = gapminder, year > 1990) %>% 
    head()
```
or (the way I like to write it).
```{r}
gapminder %>% 
    filter(year > 1990) %>% 
    head()
```
This syntax is arguably easier to read and understand than that from the base syntax. This is particularly because the tidyverse functions allow you to refer to the names of the columns within a dataframe/tibble as if they were objects in your environment. This means that you can avoid having to write something like this.
```{r}
gapminder %>% 
    filter(gapminder$year > 1990) %>% 
    head()
```
or this.
```{r}
gapminder %>% 
    filter(gapminder["year"] > 1990) %>% 
    head()
```

**TASK 13**: Use the base syntax to filter rows from the gapminder data that belong to the continent Asia and have life expectancy below 50 and year greater than 2000.

```{r}

```

**TASK 14**: Repeat the same task using `filter()`
```{r}

```

###Put it all together
Now let'd combine both row and column subsets. The code below selects all the columns other than the 4th and 6th and filters rows for which the year is greater than 1990.
```{r}
gapminder[gapminder$year > 1990, c(-4, -6)] %>% 
    head()
```

The same thing in dplyr can be done in a few ways. Using column names.
```{r}
gapminder %>% 
    filter(year > 1990) %>% 
    select(-lifeExp, -gdpPercap) %>% 
    head()
```
or (without a vector specifying col position)
```{r}
gapminder %>% 
    filter(year > 1990) %>% 
    select(-4, -6) %>% 
    head()
```
or (with a vector)
```{r}
gapminder %>% 
    filter(year > 1990) %>% 
    select(c(-4, -6)) %>% 
    head()
```

##Mutate
I will use a random sample of 10 rows of the gapminder data for this section. Let's create that first.
```{r}
gapSmall <- sample_n(gapminder, size = 10)
gapSmall
```

While we have'nt covered this in class, it is possible to add new columns of data using base R syntax. The first method uses the `$` notation
```{r}
gapSmall$yearLogical <- gapSmall$year > 1990
gapSmall
```

Another way to add columns is using `[]`
```{r}
gapSmall["hiCol"] <- "Hi I am a new column"

gapSmall
```

There are a few things to note for both these methods. Notice how I specify the name of the new column I would like to create followed use the assignment operator to assign new values to this new column.

Lets remove these two columns by only keeping those columns that share a name with the original gapminder data. 
```{r}
gapSmall <- gapSmall[names(gapminder)]

gapSmall
```

The same series of steps using dplyr syntax is as follows.
```{r}
gapSmall <- gapSmall %>% 
    mutate(yearLogical = year > 1990,
           hiCol = "Hi I am a new column")

gapSmall
```
And to remove these.
```{r}
gapSmall <- gapSmall %>% 
    select(names(gapminder))

gapSmall
```
Notice how I have to use the assignment operator to reassign the value of gapSmall each time I make changes to it.

**Task 1**: Use the base syntax to add a new column that pastes together the name of the continent and country separated by "-" to the `gapSmall` data. Give it a name of your choice.
```{r}

```

**Task 2**: Do the same thing with `dplyr` syntax.

##Summarise
Coming soon

##Group
Coming soon

##Join
Coming soon

##Visualize
Coming soon

##Reshape
Coming soon