---
title: "Class 15 - Lecture"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---

**LEARNING OBJECTIVE:** Control the flow

###Topics
1. Correction: Gather allows quasiquotation!
2. Non-vectorised conditionals (if else if else, any, all)
3. Vectorised conditionals (if_else, case_when)
4. Our first for loop.


###Basic if else statement

if (condition) {then do something} else {Do something else}

```{r}
if(TRUE) {
    "Hey there I am true"
} else {
    "I am false"
}
```

Lets feed false as the condition

```{r}
if(FALSE) {
    "Hey there I am true"
} else {
    "I am false"
}
```
So if you feed a condition which does not coerce to TRUE/FALSE it will yield an error.
```{r}
if(NA) {
    "Hey there I am true"
} else {
    "I am false"
}
```

Lets try this with an actual condition
```{r}
x <- 4
if(x == 4) {
    "Hey there I am true"
} else {
    "I am false"
}
```

```{r}
x == 4
```


Now, let's see a few examples of coercion
```{r}
if(0) {
    "Hey there I am true"
} else {
    "I am false"
}
```

0 is called a falsey value.

```{r}
0 == FALSE
```
While 1 is a truthy value.
```{r}
if(1) {
    "Hey there I am true"
} else {
    "I am false"
}
```
Any number other than zero is 'truthy'
```{r}
if(10) {
    "Hey there I am true"
} else {
    "I am false"
}
```


##IF ELSE IF ELSE

```{r}
x <- 4
if (x == 5) {
    "I am 5"
} else if(x %% 2 == 0) {
    "I am even"
} else {
    "Try again"
}
```
```{r}
x == 5
```


```{r}
x %% 2 == 0
```

```{r}
x <- 7
if (x == 5) {
    "I am 5"
} else if(x %% 2 == 0) {
    "I am even"
} else {
    "Try again"
}
```

##If else conditions are not vectorized
If else conditions evaluate a single 
```{r}
x <- 1:10

if(x %% 2 == 0) {
    "even"
} else {
    "odd"
}
```
```{r}
x
```

```{r}
x %% 2
```

```{r}
x %% 2  == 0
```

```{r}
evenNumbers <- x %% 2 == 0
evenNumbers
if(evenNumbers) {
    "even"
} else {
    "odd"
}
```

```{r}
if(c(F, T, F, T)) {
    "even"
} else {
    "odd"
}
```

DO NOT FEED AN IF STATEMENT A LOGICAL VECTOR THAT IS LONGER THAN ONE.

##Any and all

```{r}
x <- c(1, 3, 5, 7)

if(any(x %% 2 == 0)) {
    "There are even numbers"
} else {
    "Everything is odd"
}
```

```{r}
x %% 2 == 0
```

```{r}
any(x %% 2 == 0)
```

If there is a single TRUE value in a logical vector it returns TRUE.
```{r}
any(c(TRUE, FALSE, FALSE, FALSE))
```

```{r}
any(c(FALSE, FALSE, FALSE, FALSE))
```


A very practical usecase for this is when we would like to see if there are missing values in a vector.

```{r}
x <- c(2, 5, 7, 0, NA)

x
```


```{r}
is.na(x)
```

Are any values in a vector missing?
```{r}
any(is.na(x))
```



```{r}
x <- c(1, 3, 5, 7, 8, 10)

if(any(x %% 2 == 0)) {
    "There are even numbers"
} else {
    "Everything is odd"
}
```


```{r}
if(all(x %% 2 == 0)) {
    "All of them are even numbers"
} else {
    "Some of them are odd"
}
```

```{r}
all(c(T, T, T))
```

```{r}
all(c(T, F, T))
```


##Vectorized ifs

```{r}
x <- 1:10
x
```

```{r}
if(x %% 2 == 0) {
    "even"
} else {
    "odd"
}
```

The correct way to do this is as follows

```{r}
if_else(x %% 2 == 0, "even", "odd")
```

So this is very commonly used within `mutate()` in dplyr.
```{r}
library(tidyverse); library(gapminder)

gapminder %>% 
    mutate(year2000 = if_else(year >= 2000, "This are naughts", "These are the old days"))
```

What if there are multiple conditions to evaluate?
```{r}
gapminder %>% 
    mutate(year2000 = if_else(year >= 2000, 
                              "This are naughts", 
                              if_else(year >= 1980, 
                                      "still a good time", 
                                      "The really old days")))
```

The correct way to do this is using a `case_when()`
```{r}
gapminder %>% 
    mutate(year2000 = case_when(
        year >= 2000 ~ "This are naughts",
        year > 1980 ~ "still a good time",
        TRUE ~ "The really old days"))
```

This is another one of those for which the order of the conditions matter. So for instance this will not yield the correct result.

```{r}
gapminder %>% 
    mutate(year2000 = case_when(
        year > 1980 ~ "still a good time",
        year >= 2000 ~ "This are naughts",
        TRUE ~ "The really old days"))
```























