---
title: "Class 15 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: haddock
    smart: yes
    theme: readable
    df_print: paged
---


##Task Set 1: Non vectorised conditional
1. Correct the code below so that it runs.
```{r}
animal <- sample(c("Birds", "Fish"), size = 1)

if(animal = "Fish"){
    print("I can swim")
} else {
    print("I can't swim")
}
```
2. Modify the code above to check if an animal is a "Bird" and print whether it can fly or not.

3. Use an `if()` with `any()` to print "yes" if there are `any()` values greater than 4 in the vector x below.
```{r}
x <- c(3, 6, 7, 8, NA)
```

4. Modify the code above to check if there any values that are missing in x. You can tell if an element in a vector is missing using the function `is.na()`.

5. Modify the code you wrote for task 3 to check if there are any values greater than 8 instead of checking if there are any missing values. This will likely throw an error. Can you reason why this might be giving you an error? See the documentation for `any()` to see how to address this.

6. Check if `all()` the values in vector x are not missing. If yes print "yes" else print "no".


##Task Set 5: Vectorized Ifs
1. Use an `if_else()` statement to create a character vector that indicates whether a given value in vector x (from previous task set) is missing. It should be marked "missing" if a value is missing and "not missing" otherwise. 

2. Load the gapminder and tidyverse packages

3. Use the `if_else()` in a `mutate()` to add a new column to the gapminder data that indicates whether the year is greater than 2000.

4. Make sure you have already loaded the tidyverse and gapminder before you start this. Complete the code below (the `group_by()` and `if_else()`). For every year I would like to mark the `incomeGroup` variable as "Higher" if it is above the median `gdpPercap` for that year (global median) and "Lower" otherwise. The subsequent plot explores the relationship between `incomeGroup` and the `lifeExp`. How would you interpret it?
```{r, fig.width = 9}
gapminder %>% 
    group_by() %>% 
    mutate(incomeGroup = if_else()) %>% 
    ggplot(aes(x = as.factor(year), y = lifeExp, colour = incomeGroup)) +
    geom_boxplot(outlier.alpha = 0.3) +
    labs(title = str_wrap("Over time life expectancy in high income countries have converged while those in low income countries have diverged", width = 70),
         y = "Life Expectancy",
         colour = "Income Group") +
    ggthemes::theme_tufte() +
    theme(
        legend.position = "bottom",
        axis.title.x = element_blank()
    )
    
    
```

4. Complete the following code to create a new variable called incomeGroup that cuts the `gdpPercap` variable based on the median value of the `gdpPercap` every year. The final variable should have 4 levels - Low income, low middle income, middle income, high middle income and high income. I have complete one of the levels for you in the code. Complete the rest - gdpPercap < 0.75 x median(gdpPercap) is "Low middle income", gdpPercap < 2 x median(gdpPercap) is "Middle income", gdpPercap < 4 x median(gdpPercap) is "High middle income" and everything else is high income.
```{r, fig.width = 12}
gapminder %>% 
    group_by(year) %>% 
    mutate(incomeGroups = case_when(
        gdpPercap < 0.25 * median(gdpPercap) ~ "Low income",
        gdpPercap <  ~ "Low middle income",
         ~ "Middle income",
         ~ "High middle income",
        TRUE ~ "High Income"
    )) %>% 
    ungroup() %>% 
    ggplot(aes(x = as.factor(year), y = lifeExp)) +
    geom_boxplot() +
    facet_wrap(~incomeGroups, nrow = 1) +
    labs(title = str_wrap("Over time life expectancy in high income countries have converged while those in low income countries have diverged", width = 90),
         y = "Life Expectancy",
         colour = "Income Group") +
    theme_tufte() +
    theme(
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45)
    )
```


##Task Set 3: Write a for loop
1. Create a sequence of numbers from 1 to 50 store it in a variable

2. Write a for loop that returns "even" if a number in the vector you created above is even and "odd" otherwise. Store the results from this in a variable called odds_evens. You can initiate the `odds_evens` vector like this - `odds_evens <- vector("character", length = length(x))`. You can establish if a value is even by checking if the remainder when you divide it by 2 is zero. The syntax for this is `value %% 2 == 0`, where value is the number you are comparing.

3. Now do the same thing with an `if_else()` statement and store the results in a 




